---
title: "Comparing models with different genetics parameters"
author: "Sarah P. Flanagan"
date:  "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::pdf_document2:
    fig_caption: yes
    keep_tex: yes
    number_sections: no
    template: manuscript.latex
    toc: yes
    toc_depth: 2
spacing: singlespacing
fontsize: 11pt
capsize: normalsize
documentclass: article
footerdate: yes
graphics: yes
csl: evolution.csl
editor_options:
  chunk_output_type: console
---

The purpose of this model is to understand how genetic architectures of alternative reproductive tactics impact their maintenance in populations. I'm using an individual-based simulation model with different selection scenarios, types of alternative tactics, and genetic architectures (genome-wide additive genetic variance vs supergenes). To explore the parameter space, I'm also using a mathematical model without the genetic architectures to identify regions of parameter space where we expect multiple morphs to be maintained. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,out.extra='',fig.pos="H",dpi=200,fig.height = 7,fig.width = 7)
knitr::opts_knit$set(root.dir='../results/') #change
```
```{r opts, echo = FALSE}
library(knitr)
knitr::opts_chunk$set(
  fig.path = "../figs/"
)
```
```{r librarysetup, echo=FALSE,echo=FALSE,message=FALSE}
library(RColorBrewer)
library(scales)
library(sm)
library(kableExtra)
library(plotly)
library(dplyr)
library(vegan)
library(tidyr)
source("../R/002_freq_functions.R")
tmp<-sapply(list.files(path="../../spfTools/R/",full.names = TRUE),source)
source("../morph_predictions/check_freqs.R")
source("../morph_predictions/morph_gens_ns.R")

cols<-c(courter="#33a02c",parent="#1f78b4",cpref="#b2df8a",ppref="#a6cee3")
cols2<-c(CP='#d7191c',NCP='#fdae61',CNP='#abd9e9',NCNP='#2c7bb6')
```
```{r baseCols}
base_keep_cols<- c("Pop","PopSize","NumMal","NumFem","ParentThresh","ParentFreq","CourterThresh","CourterFreq","FreqNcNp","FreqCNp",
                   "FreqNcP","FreqCP","Courter2NonRS","Parent2NonSurvival")
poly_keep_cols<- c("Pop","PopSize","NumMal","NumFem","ParentThresh","ParentFreq","CourterThresh","CourterFreq","FreqNcNp","FreqCNp",
                    "FreqNcP","FreqCP","Polygyny")

```



## Low diversity parameter space

The [results from the mathematical model](https://spflanagan.shinyapps.io/morph_predictions/) were inspected to identify parameter regions where multiple morphs were *not* maintained. I focused on parameter sets when morphs had equal initial frequencies, as the simulation model initiates populations with nearly equal frequencies of the morphs. I selected the low diversity parameters in Table \@ref(tab:lowDivParams) to implement in the simulation model with a single locus underpinning the courter and parent traits. 

```{r lowDivParams}
highDivParams<-data.frame(
  parameter = c("r (in shiny app)",
                "r_CP",
                "r_CN",
                "r_NP",
                "r_NN",
                "c",
                "number of sneakers"),
  value = c( 2, 
             8,
             8,
             4,
             4,
             0.5,
             2)
)
kable(highDivParams, "latex",booktab=TRUE,
      caption="Table of parameters selected to create a low-diversity population. All morphs started at equal frequencies.") %>%
  kable_styling(latex_options = "HOLD_position")

```

### QTLs

```{r lowDivGenetics}
lowDiv<-get.morph.freqs(plot.pc.reps(pattern="lowDiversity.*monogamy_nm.*summary.txt",
                                     path="/mnt/BigData/fixedART-results/qtls",
                                     cols,make.plot=FALSE))
lnames<-rownames(lowDiv)
```

```{r}
qs<-unique(gsub(".*lowDiversity_.*_q(\\d+)_c(\\d)_\\d_summary.txt_\\d","\\1",lnames))
cs<-unique(gsub(".*lowDiversity_.*_q(\\d+)_c(\\d)_\\d_summary.txt_\\d","\\1",lnames))
```

```{r lowDivQTLs16, fig.cap="Final generation frequencies based on the model with 16 QTLs, with parameters that are expected to result in no diversity based on the mathematical model. In 'monogamy' cases, the males were constrained to only accept eggs from one female into their nests, whereas under 'polygyny' males could mate with multiple females. In both cases females mated once. Parameter settings lablled with 'nm' were runs where females did not mate if they could not find a suitable mate (i.e., if no courters were available)."}

var_patts<-unique(gsub(".*lowDiversity_(.*)_\\d_summary.txt_\\d","\\1",
                       lnames[grep(paste0("q",qs[1]),lnames)]))


par(mfrow=c(2,ceiling(length(var_patts)/2)),oma=c(3,1,4,1),mar=c(3,1,4,1),xpd=TRUE)
varDat<-lapply(var_patts,function(pattern){
  dat<-lowDiv[grep(paste0(pattern,"_\\d"),rownames(lowDiv)),]
  bp<-barplot(t(as.matrix(dat[,c("FreqNcNp", "FreqCNp", "FreqNcP", "FreqCP")])),
        col=cols2[c("NCNP","CNP","NCP","CP")],
        names.arg =gsub("^.*(\\d)_summary.txt.*$","\\1",rownames(dat)),
        las=2,
        border=NA,
        main=gsub("qtls_monogamy_nm_(.*)","\\1",pattern))
  text(x=bp,y=0.5,srt=90,labels = dat$FreqCP)

  return(dat)
})
par(fig=c(0, 1, 0, 1), oma=c(0, 0, 0, 0), mar=c(0, 0, 0, 0), new=TRUE)
plot(0,0, type='n', bty='n', xaxt='n', yaxt='n')
legend("top",bty='n',legend = c("Courter/Parent","Courter/Non-parent","Non-courter/Parent","Non-courter/Non-parent"),
       col=cols2[c("CP","CNP","NCP","NCNP")],pch=15,xpd = TRUE,ncol=2)

```

```{r lowDivQTLs32, fig.cap="Final generation frequencies based on the model with 32 QTLs, with parameters that are expected to result in no diversity based on the mathematical model. In 'monogamy' cases, the males were constrained to only accept eggs from one female into their nests, whereas under 'polygyny' males could mate with multiple females. In both cases females mated once. Parameter settings lablled with 'nm' were runs where females did not mate if they could not find a suitable mate (i.e., if no courters were available)."}

var_patts<-unique(gsub(".*lowDiversity_(.*)_\\d_summary.txt_\\d","\\1",
                       lnames[grep(paste0("q",qs[2]),lnames)]))


par(mfrow=c(2,ceiling(length(var_patts)/2)),oma=c(3,1,4,1),mar=c(3,1,4,1),xpd=TRUE)
varDat<-lapply(var_patts,function(pattern){
  dat<-lowDiv[grep(paste0(pattern,"_\\d"),rownames(lowDiv)),]
  bp<-barplot(t(as.matrix(dat[,c("FreqNcNp", "FreqCNp", "FreqNcP", "FreqCP")])),
        col=cols2[c("NCNP","CNP","NCP","CP")],
        names.arg =gsub("^.*(\\d)_summary.txt.*$","\\1",rownames(dat)),
        las=2,
        border=NA,
        main=gsub("qtls_monogamy_nm_(.*)","\\1",pattern))
  text(x=bp,y=0.5,srt=90,labels = dat$FreqCP)

  return(dat)
})
par(fig=c(0, 1, 0, 1), oma=c(0, 0, 0, 0), mar=c(0, 0, 0, 0), new=TRUE)
plot(0,0, type='n', bty='n', xaxt='n', yaxt='n')
legend("top",bty='n',legend = c("Courter/Parent","Courter/Non-parent","Non-courter/Parent","Non-courter/Non-parent"),
       col=cols2[c("CP","CNP","NCP","NCNP")],pch=15,xpd = TRUE,ncol=2)

```

```{r lowDivQTLs64, fig.cap="Final generation frequencies based on the model with 64 QTLs, with parameters that are expected to result in no diversity based on the mathematical model. In 'monogamy' cases, the males were constrained to only accept eggs from one female into their nests, whereas under 'polygyny' males could mate with multiple females. In both cases females mated once. Parameter settings lablled with 'nm' were runs where females did not mate if they could not find a suitable mate (i.e., if no courters were available)."}

var_patts<-unique(gsub(".*lowDiversity_(.*)_\\d_summary.txt_\\d","\\1",
                       lnames[grep(paste0("q",qs[3]),lnames)]))


par(mfrow=c(2,ceiling(length(var_patts)/2)),oma=c(3,1,4,1),mar=c(3,1,4,1),xpd=TRUE)
varDat<-lapply(var_patts,function(pattern){
  dat<-lowDiv[grep(paste0(pattern,"_\\d"),rownames(lowDiv)),]
  bp<-barplot(t(as.matrix(dat[,c("FreqNcNp", "FreqCNp", "FreqNcP", "FreqCP")])),
        col=cols2[c("NCNP","CNP","NCP","CP")],
        names.arg =gsub("^.*(\\d)_summary.txt.*$","\\1",rownames(dat)),
        las=2,
        border=NA,
        main=gsub("qtls_monogamy_nm_(.*)","\\1",pattern))
  text(x=bp,y=0.5,srt=90,labels = dat$FreqCP)

  return(dat)
})
par(fig=c(0, 1, 0, 1), oma=c(0, 0, 0, 0), mar=c(0, 0, 0, 0), new=TRUE)
plot(0,0, type='n', bty='n', xaxt='n', yaxt='n')
legend("top",bty='n',legend = c("Courter/Parent","Courter/Non-parent","Non-courter/Parent","Non-courter/Non-parent"),
       col=cols2[c("CP","CNP","NCP","NCNP")],pch=15,xpd = TRUE,ncol=2)

```

```{r lowDivQTLs8, fig.cap="Final generation frequencies based on the model with 8 QTLs, with parameters that are expected to result in no diversity based on the mathematical model. In 'monogamy' cases, the males were constrained to only accept eggs from one female into their nests, whereas under 'polygyny' males could mate with multiple females. In both cases females mated once. Parameter settings lablled with 'nm' were runs where females did not mate if they could not find a suitable mate (i.e., if no courters were available)."}

var_patts<-unique(gsub(".*lowDiversity_(.*)_\\d_summary.txt_\\d","\\1",
                       lnames[grep(paste0("q",qs[4]),lnames)]))


par(mfrow=c(2,ceiling(length(var_patts)/2)),oma=c(3,1,4,1),mar=c(3,1,4,1),xpd=TRUE)
varDat<-lapply(var_patts,function(pattern){
  dat<-lowDiv[grep(paste0(pattern,"_\\d"),rownames(lowDiv)),]
  bp<-barplot(t(as.matrix(dat[,c("FreqNcNp", "FreqCNp", "FreqNcP", "FreqCP")])),
        col=cols2[c("NCNP","CNP","NCP","CP")],
        names.arg =gsub("^.*(\\d)_summary.txt.*$","\\1",rownames(dat)),
        las=2,
        border=NA,
        main=gsub("qtls_monogamy_nm_(.*)","\\1",pattern))
  text(x=bp,y=0.5,srt=90,labels = dat$FreqCP)

  return(dat)
})
par(fig=c(0, 1, 0, 1), oma=c(0, 0, 0, 0), mar=c(0, 0, 0, 0), new=TRUE)
plot(0,0, type='n', bty='n', xaxt='n', yaxt='n')
legend("top",bty='n',legend = c("Courter/Parent","Courter/Non-parent","Non-courter/Parent","Non-courter/Non-parent"),
       col=cols2[c("CP","CNP","NCP","NCNP")],pch=15,xpd = TRUE,ncol=2)

```

### Supergenes


```{r lowDivGenetics2}
lowDiv<-get.morph.freqs(plot.pc.reps(pattern="lowDiversity.*monogamy_nm.*summary.txt",
                                     path="/mnt/BigData/fixedART-results/supergene",
                                     cols,make.plot=FALSE))
lnames<-rownames(lowDiv)
```

```{r}
qs<-unique(gsub(".*lowDiversity_.*_q(\\d+)_c(\\d)_\\d_summary.txt_\\d","\\1",lnames))
cs<-unique(gsub(".*lowDiversity_.*_q(\\d+)_c(\\d)_\\d_summary.txt_\\d","\\1",lnames))
```

```{r lowDivQTLs16super, fig.cap="Final generation frequencies based on the model with 16 QTLs in a supergene, with parameters that are expected to result in no diversity based on the mathematical model. In 'monogamy' cases, the males were constrained to only accept eggs from one female into their nests, whereas under 'polygyny' males could mate with multiple females. In both cases females mated once. Parameter settings lablled with 'nm' were runs where females did not mate if they could not find a suitable mate (i.e., if no courters were available)."}

var_patts<-unique(gsub(".*lowDiversity_(.*)_\\d_summary.txt_\\d","\\1",
                       lnames[grep(paste0("q",qs[1]),lnames)]))


par(mfrow=c(2,ceiling(length(var_patts)/2)),oma=c(3,1,4,1),mar=c(3,1,4,1),xpd=TRUE)
varDat<-lapply(var_patts,function(pattern){
  dat<-lowDiv[grep(paste0(pattern,"_\\d"),rownames(lowDiv)),]
  bp<-barplot(t(as.matrix(dat[,c("FreqNcNp", "FreqCNp", "FreqNcP", "FreqCP")])),
        col=cols2[c("NCNP","CNP","NCP","CP")],
        names.arg =gsub("^.*(\\d)_summary.txt.*$","\\1",rownames(dat)),
        las=2,
        border=NA,
        main=gsub("qtls_monogamy_nm_(.*)","\\1",pattern))
  text(x=bp,y=0.5,srt=90,labels = dat$FreqCP)

  return(dat)
})
par(fig=c(0, 1, 0, 1), oma=c(0, 0, 0, 0), mar=c(0, 0, 0, 0), new=TRUE)
plot(0,0, type='n', bty='n', xaxt='n', yaxt='n')
legend("top",bty='n',legend = c("Courter/Parent","Courter/Non-parent","Non-courter/Parent","Non-courter/Non-parent"),
       col=cols2[c("CP","CNP","NCP","NCNP")],pch=15,xpd = TRUE,ncol=2)

```

```{r lowDivQTLs32super, fig.cap="Final generation frequencies based on the model with 32 QTLs in a supergene, with parameters that are expected to result in no diversity based on the mathematical model. In 'monogamy' cases, the males were constrained to only accept eggs from one female into their nests, whereas under 'polygyny' males could mate with multiple females. In both cases females mated once. Parameter settings lablled with 'nm' were runs where females did not mate if they could not find a suitable mate (i.e., if no courters were available)."}

var_patts<-unique(gsub(".*lowDiversity_(.*)_\\d_summary.txt_\\d","\\1",
                       lnames[grep(paste0("q",qs[2]),lnames)]))


par(mfrow=c(2,ceiling(length(var_patts)/2)),oma=c(3,1,4,1),mar=c(3,1,4,1),xpd=TRUE)
varDat<-lapply(var_patts,function(pattern){
  dat<-lowDiv[grep(paste0(pattern,"_\\d"),rownames(lowDiv)),]
  bp<-barplot(t(as.matrix(dat[,c("FreqNcNp", "FreqCNp", "FreqNcP", "FreqCP")])),
        col=cols2[c("NCNP","CNP","NCP","CP")],
        names.arg =gsub("^.*(\\d)_summary.txt.*$","\\1",rownames(dat)),
        las=2,
        border=NA,
        main=gsub("qtls_monogamy_nm_(.*)","\\1",pattern))
  text(x=bp,y=0.5,srt=90,labels = dat$FreqCP)

  return(dat)
})
par(fig=c(0, 1, 0, 1), oma=c(0, 0, 0, 0), mar=c(0, 0, 0, 0), new=TRUE)
plot(0,0, type='n', bty='n', xaxt='n', yaxt='n')
legend("top",bty='n',legend = c("Courter/Parent","Courter/Non-parent","Non-courter/Parent","Non-courter/Non-parent"),
       col=cols2[c("CP","CNP","NCP","NCNP")],pch=15,xpd = TRUE,ncol=2)

```

```{r lowDivQTLs64super, fig.cap="Final generation frequencies based on the model with 64 QTLs in a supergene, with parameters that are expected to result in no diversity based on the mathematical model. In 'monogamy' cases, the males were constrained to only accept eggs from one female into their nests, whereas under 'polygyny' males could mate with multiple females. In both cases females mated once. Parameter settings lablled with 'nm' were runs where females did not mate if they could not find a suitable mate (i.e., if no courters were available)."}

var_patts<-unique(gsub(".*lowDiversity_(.*)_\\d_summary.txt_\\d","\\1",
                       lnames[grep(paste0("q",qs[3]),lnames)]))


par(mfrow=c(2,ceiling(length(var_patts)/2)),oma=c(3,1,4,1),mar=c(3,1,4,1),xpd=TRUE)
varDat<-lapply(var_patts,function(pattern){
  dat<-lowDiv[grep(paste0(pattern,"_\\d"),rownames(lowDiv)),]
  bp<-barplot(t(as.matrix(dat[,c("FreqNcNp", "FreqCNp", "FreqNcP", "FreqCP")])),
        col=cols2[c("NCNP","CNP","NCP","CP")],
        names.arg =gsub("^.*(\\d)_summary.txt.*$","\\1",rownames(dat)),
        las=2,
        border=NA,
        main=gsub("qtls_monogamy_nm_(.*)","\\1",pattern))
  text(x=bp,y=0.5,srt=90,labels = dat$FreqCP)

  return(dat)
})
par(fig=c(0, 1, 0, 1), oma=c(0, 0, 0, 0), mar=c(0, 0, 0, 0), new=TRUE)
plot(0,0, type='n', bty='n', xaxt='n', yaxt='n')
legend("top",bty='n',legend = c("Courter/Parent","Courter/Non-parent","Non-courter/Parent","Non-courter/Non-parent"),
       col=cols2[c("CP","CNP","NCP","NCNP")],pch=15,xpd = TRUE,ncol=2)

```

```{r lowDivQTLs8 super, fig.cap="Final generation frequencies based on the model with 8 QTLs in a supergene, with parameters that are expected to result in no diversity based on the mathematical model. In 'monogamy' cases, the males were constrained to only accept eggs from one female into their nests, whereas under 'polygyny' males could mate with multiple females. In both cases females mated once. Parameter settings lablled with 'nm' were runs where females did not mate if they could not find a suitable mate (i.e., if no courters were available)."}

var_patts<-unique(gsub(".*lowDiversity_(.*)_\\d_summary.txt_\\d","\\1",
                       lnames[grep(paste0("q",qs[4]),lnames)]))


par(mfrow=c(2,ceiling(length(var_patts)/2)),oma=c(3,1,4,1),mar=c(3,1,4,1),xpd=TRUE)
varDat<-lapply(var_patts,function(pattern){
  dat<-lowDiv[grep(paste0(pattern,"_\\d"),rownames(lowDiv)),]
  bp<-barplot(t(as.matrix(dat[,c("FreqNcNp", "FreqCNp", "FreqNcP", "FreqCP")])),
        col=cols2[c("NCNP","CNP","NCP","CP")],
        names.arg =gsub("^.*(\\d)_summary.txt.*$","\\1",rownames(dat)),
        las=2,
        border=NA,
        main=gsub("qtls_monogamy_nm_(.*)","\\1",pattern))
  text(x=bp,y=0.5,srt=90,labels = dat$FreqCP)

  return(dat)
})
par(fig=c(0, 1, 0, 1), oma=c(0, 0, 0, 0), mar=c(0, 0, 0, 0), new=TRUE)
plot(0,0, type='n', bty='n', xaxt='n', yaxt='n')
legend("top",bty='n',legend = c("Courter/Parent","Courter/Non-parent","Non-courter/Parent","Non-courter/Non-parent"),
       col=cols2[c("CP","CNP","NCP","NCNP")],pch=15,xpd = TRUE,ncol=2)

```


## High diversity parameter space

The [results from the mathematical model](https://spflanagan.shinyapps.io/morph_predictions/) were inspected to identify parameter regions where multiple morphs were maintained. I focused on parameter sets when morphs had equal initial frequencies, as the simulation model initiates populations with nearly equal frequencies of the morphs. I selected the high-diversity parameters in Table \@ref(tab:highDivParams) to implement in the simulation model with a single locus underpinning the courter and parent traits. 

```{r highDivParams}
highDivParams<-data.frame(
  parameter = c("r (in shiny app)",
                "r_CP",
                "r_CN",
                "r_NP",
                "r_NN",
                "c",
                "number of sneakers"),
  value = c( 0.75, 
             6,
             6,
             8,
             8,
             0.7,
             2)
)
kable(highDivParams, "latex",booktab=TRUE,
      caption="Table of parameters selected to create a high-diversity population. All morphs started at equal frequencies.") %>%
  kable_styling(latex_options = "HOLD_position")

```

### QTLs

```{r highDivGenetics}
lowDiv<-get.morph.freqs(plot.pc.reps(pattern="highDiversity.*monogamy_nm.*summary.txt",
                                     path="/mnt/BigData/fixedART-results/qtls",
                                     cols,make.plot=FALSE))
lnames<-rownames(lowDiv)
```

```{r}
qs<-unique(gsub(".*highDiversity_.*_q(\\d+)_c(\\d)_\\d_summary.txt_\\d","\\1",lnames))
cs<-unique(gsub(".*highDiversity_.*_q(\\d+)_c(\\d)_\\d_summary.txt_\\d","\\1",lnames))
```

```{r highDivQTLs16, fig.cap="Final generation frequencies based on the model with 16 QTLs, with parameters that are expected to result in high diversity based on the mathematical model. In 'monogamy' cases, the males were constrained to only accept eggs from one female into their nests, whereas under 'polygyny' males could mate with multiple females. In both cases females mated once. Parameter settings lablled with 'nm' were runs where females did not mate if they could not find a suitable mate (i.e., if no courters were available)."}

var_patts<-unique(gsub(".*highDiversity_(.*)_\\d_summary.txt_\\d","\\1",
                       lnames[grep(paste0("q",qs[1]),lnames)]))


par(mfrow=c(2,ceiling(length(var_patts)/2)),oma=c(3,1,4,1),mar=c(3,1,4,1),xpd=TRUE)
varDat<-lapply(var_patts,function(pattern){
  dat<-lowDiv[grep(paste0(pattern,"_\\d"),rownames(lowDiv)),]
  bp<-barplot(t(as.matrix(dat[,c("FreqNcNp", "FreqCNp", "FreqNcP", "FreqCP")])),
        col=cols2[c("NCNP","CNP","NCP","CP")],
        names.arg =gsub("^.*(\\d)_summary.txt.*$","\\1",rownames(dat)),
        las=2,
        border=NA,
        main=gsub("qtls_monogamy_nm_(.*)","\\1",pattern))
  text(x=bp,y=0.5,srt=90,labels = dat$FreqCP)

  return(dat)
})
par(fig=c(0, 1, 0, 1), oma=c(0, 0, 0, 0), mar=c(0, 0, 0, 0), new=TRUE)
plot(0,0, type='n', bty='n', xaxt='n', yaxt='n')
legend("top",bty='n',legend = c("Courter/Parent","Courter/Non-parent","Non-courter/Parent","Non-courter/Non-parent"),
       col=cols2[c("CP","CNP","NCP","NCNP")],pch=15,xpd = TRUE,ncol=2)

```

```{r highDivQTLs32, fig.cap="Final generation frequencies based on the model with 32 QTLs, with parameters that are expected to result in high diversity based on the mathematical model. In 'monogamy' cases, the males were constrained to only accept eggs from one female into their nests, whereas under 'polygyny' males could mate with multiple females. In both cases females mated once. Parameter settings lablled with 'nm' were runs where females did not mate if they could not find a suitable mate (i.e., if no courters were available)."}

var_patts<-unique(gsub(".*highDiversity_(.*)_\\d_summary.txt_\\d","\\1",
                       lnames[grep(paste0("q",qs[2]),lnames)]))


par(mfrow=c(2,ceiling(length(var_patts)/2)),oma=c(3,1,4,1),mar=c(3,1,4,1),xpd=TRUE)
varDat<-lapply(var_patts,function(pattern){
  dat<-lowDiv[grep(paste0(pattern,"_\\d"),rownames(lowDiv)),]
  bp<-barplot(t(as.matrix(dat[,c("FreqNcNp", "FreqCNp", "FreqNcP", "FreqCP")])),
        col=cols2[c("NCNP","CNP","NCP","CP")],
        names.arg =gsub("^.*(\\d)_summary.txt.*$","\\1",rownames(dat)),
        las=2,
        border=NA,
        main=gsub("qtls_monogamy_nm_(.*)","\\1",pattern))
  text(x=bp,y=0.5,srt=90,labels = dat$FreqCP)

  return(dat)
})
par(fig=c(0, 1, 0, 1), oma=c(0, 0, 0, 0), mar=c(0, 0, 0, 0), new=TRUE)
plot(0,0, type='n', bty='n', xaxt='n', yaxt='n')
legend("top",bty='n',legend = c("Courter/Parent","Courter/Non-parent","Non-courter/Parent","Non-courter/Non-parent"),
       col=cols2[c("CP","CNP","NCP","NCNP")],pch=15,xpd = TRUE,ncol=2)

```

```{r highDivQTLs64, fig.cap="Final generation frequencies based on the model with 64 QTLs, with parameters that are expected to result in high diversity based on the mathematical model. In 'monogamy' cases, the males were constrained to only accept eggs from one female into their nests, whereas under 'polygyny' males could mate with multiple females. In both cases females mated once. Parameter settings lablled with 'nm' were runs where females did not mate if they could not find a suitable mate (i.e., if no courters were available)."}

var_patts<-unique(gsub(".*highDiversity_(.*)_\\d_summary.txt_\\d","\\1",
                       lnames[grep(paste0("q",qs[3]),lnames)]))


par(mfrow=c(2,ceiling(length(var_patts)/2)),oma=c(3,1,4,1),mar=c(3,1,4,1),xpd=TRUE)
varDat<-lapply(var_patts,function(pattern){
  dat<-lowDiv[grep(paste0(pattern,"_\\d"),rownames(lowDiv)),]
  bp<-barplot(t(as.matrix(dat[,c("FreqNcNp", "FreqCNp", "FreqNcP", "FreqCP")])),
        col=cols2[c("NCNP","CNP","NCP","CP")],
        names.arg =gsub("^.*(\\d)_summary.txt.*$","\\1",rownames(dat)),
        las=2,
        border=NA,
        main=gsub("qtls_monogamy_nm_(.*)","\\1",pattern))
  text(x=bp,y=0.5,srt=90,labels = dat$FreqCP)

  return(dat)
})
par(fig=c(0, 1, 0, 1), oma=c(0, 0, 0, 0), mar=c(0, 0, 0, 0), new=TRUE)
plot(0,0, type='n', bty='n', xaxt='n', yaxt='n')
legend("top",bty='n',legend = c("Courter/Parent","Courter/Non-parent","Non-courter/Parent","Non-courter/Non-parent"),
       col=cols2[c("CP","CNP","NCP","NCNP")],pch=15,xpd = TRUE,ncol=2)

```

```{r highDivQTLs8, fig.cap="Final generation frequencies based on the model with 8 QTLs, with parameters that are expected to result in high diversity based on the mathematical model. In 'monogamy' cases, the males were constrained to only accept eggs from one female into their nests, whereas under 'polygyny' males could mate with multiple females. In both cases females mated once. Parameter settings lablled with 'nm' were runs where females did not mate if they could not find a suitable mate (i.e., if no courters were available)."}

var_patts<-unique(gsub(".*highDiversity_(.*)_\\d_summary.txt_\\d","\\1",
                       lnames[grep(paste0("q",qs[4]),lnames)]))


par(mfrow=c(2,ceiling(length(var_patts)/2)),oma=c(3,1,4,1),mar=c(3,1,4,1),xpd=TRUE)
varDat<-lapply(var_patts,function(pattern){
  dat<-lowDiv[grep(paste0(pattern,"_\\d"),rownames(lowDiv)),]
  bp<-barplot(t(as.matrix(dat[,c("FreqNcNp", "FreqCNp", "FreqNcP", "FreqCP")])),
        col=cols2[c("NCNP","CNP","NCP","CP")],
        names.arg =gsub("^.*(\\d)_summary.txt.*$","\\1",rownames(dat)),
        las=2,
        border=NA,
        main=gsub("qtls_monogamy_nm_(.*)","\\1",pattern))
  text(x=bp,y=0.5,srt=90,labels = dat$FreqCP)

  return(dat)
})
par(fig=c(0, 1, 0, 1), oma=c(0, 0, 0, 0), mar=c(0, 0, 0, 0), new=TRUE)
plot(0,0, type='n', bty='n', xaxt='n', yaxt='n')
legend("top",bty='n',legend = c("Courter/Parent","Courter/Non-parent","Non-courter/Parent","Non-courter/Non-parent"),
       col=cols2[c("CP","CNP","NCP","NCNP")],pch=15,xpd = TRUE,ncol=2)

```

### Supergenes


```{r highDivGenetics2}
lowDiv<-get.morph.freqs(plot.pc.reps(pattern="highDiversity.*monogamy_nm.*summary.txt",
                                     path="/mnt/BigData/fixedART-results/supergene",
                                     cols,make.plot=FALSE))
lnames<-rownames(lowDiv)
```

```{r}
qs<-unique(gsub(".*highDiversity_.*_q(\\d+)_c(\\d)_\\d_summary.txt_\\d","\\1",lnames))
cs<-unique(gsub(".*highDiversity_.*_q(\\d+)_c(\\d)_\\d_summary.txt_\\d","\\1",lnames))
```

```{r highDivQTLs16super, fig.cap="Final generation frequencies based on the model with 16 QTLs in a supergene, with parameters that are expected to result in high diversity based on the mathematical model. In 'monogamy' cases, the males were constrained to only accept eggs from one female into their nests, whereas under 'polygyny' males could mate with multiple females. In both cases females mated once. Parameter settings lablled with 'nm' were runs where females did not mate if they could not find a suitable mate (i.e., if no courters were available)."}

var_patts<-unique(gsub(".*highDiversity_(.*)_\\d_summary.txt_\\d","\\1",
                       lnames[grep(paste0("q",qs[1]),lnames)]))


par(mfrow=c(2,ceiling(length(var_patts)/2)),oma=c(3,1,4,1),mar=c(3,1,4,1),xpd=TRUE)
varDat<-lapply(var_patts,function(pattern){
  dat<-lowDiv[grep(paste0(pattern,"_\\d"),rownames(lowDiv)),]
  bp<-barplot(t(as.matrix(dat[,c("FreqNcNp", "FreqCNp", "FreqNcP", "FreqCP")])),
        col=cols2[c("NCNP","CNP","NCP","CP")],
        names.arg =gsub("^.*(\\d)_summary.txt.*$","\\1",rownames(dat)),
        las=2,
        border=NA,
        main=gsub("qtls_monogamy_nm_(.*)","\\1",pattern))
  text(x=bp,y=0.5,srt=90,labels = dat$FreqCP)

  return(dat)
})
par(fig=c(0, 1, 0, 1), oma=c(0, 0, 0, 0), mar=c(0, 0, 0, 0), new=TRUE)
plot(0,0, type='n', bty='n', xaxt='n', yaxt='n')
legend("top",bty='n',legend = c("Courter/Parent","Courter/Non-parent","Non-courter/Parent","Non-courter/Non-parent"),
       col=cols2[c("CP","CNP","NCP","NCNP")],pch=15,xpd = TRUE,ncol=2)

```

```{r highDivQTLs32super, fig.cap="Final generation frequencies based on the model with 32 QTLs in a supergene, with parameters that are expected to result in high diversity based on the mathematical model. In 'monogamy' cases, the males were constrained to only accept eggs from one female into their nests, whereas under 'polygyny' males could mate with multiple females. In both cases females mated once. Parameter settings lablled with 'nm' were runs where females did not mate if they could not find a suitable mate (i.e., if no courters were available)."}

var_patts<-unique(gsub(".*highDiversity_(.*)_\\d_summary.txt_\\d","\\1",
                       lnames[grep(paste0("q",qs[2]),lnames)]))


par(mfrow=c(2,ceiling(length(var_patts)/2)),oma=c(3,1,4,1),mar=c(3,1,4,1),xpd=TRUE)
varDat<-lapply(var_patts,function(pattern){
  dat<-lowDiv[grep(paste0(pattern,"_\\d"),rownames(lowDiv)),]
  bp<-barplot(t(as.matrix(dat[,c("FreqNcNp", "FreqCNp", "FreqNcP", "FreqCP")])),
        col=cols2[c("NCNP","CNP","NCP","CP")],
        names.arg =gsub("^.*(\\d)_summary.txt.*$","\\1",rownames(dat)),
        las=2,
        border=NA,
        main=gsub("qtls_monogamy_nm_(.*)","\\1",pattern))
  text(x=bp,y=0.5,srt=90,labels = dat$FreqCP)

  return(dat)
})
par(fig=c(0, 1, 0, 1), oma=c(0, 0, 0, 0), mar=c(0, 0, 0, 0), new=TRUE)
plot(0,0, type='n', bty='n', xaxt='n', yaxt='n')
legend("top",bty='n',legend = c("Courter/Parent","Courter/Non-parent","Non-courter/Parent","Non-courter/Non-parent"),
       col=cols2[c("CP","CNP","NCP","NCNP")],pch=15,xpd = TRUE,ncol=2)

```

```{r highDivQTLs64super, fig.cap="Final generation frequencies based on the model with 64 QTLs in a supergene, with parameters that are expected to result in high diversity based on the mathematical model. In 'monogamy' cases, the males were constrained to only accept eggs from one female into their nests, whereas under 'polygyny' males could mate with multiple females. In both cases females mated once. Parameter settings lablled with 'nm' were runs where females did not mate if they could not find a suitable mate (i.e., if no courters were available)."}

var_patts<-unique(gsub(".*highDiversity_(.*)_\\d_summary.txt_\\d","\\1",
                       lnames[grep(paste0("q",qs[3]),lnames)]))


par(mfrow=c(2,ceiling(length(var_patts)/2)),oma=c(3,1,4,1),mar=c(3,1,4,1),xpd=TRUE)
varDat<-lapply(var_patts,function(pattern){
  dat<-lowDiv[grep(paste0(pattern,"_\\d"),rownames(lowDiv)),]
  bp<-barplot(t(as.matrix(dat[,c("FreqNcNp", "FreqCNp", "FreqNcP", "FreqCP")])),
        col=cols2[c("NCNP","CNP","NCP","CP")],
        names.arg =gsub("^.*(\\d)_summary.txt.*$","\\1",rownames(dat)),
        las=2,
        border=NA,
        main=gsub("qtls_monogamy_nm_(.*)","\\1",pattern))
  text(x=bp,y=0.5,srt=90,labels = dat$FreqCP)

  return(dat)
})
par(fig=c(0, 1, 0, 1), oma=c(0, 0, 0, 0), mar=c(0, 0, 0, 0), new=TRUE)
plot(0,0, type='n', bty='n', xaxt='n', yaxt='n')
legend("top",bty='n',legend = c("Courter/Parent","Courter/Non-parent","Non-courter/Parent","Non-courter/Non-parent"),
       col=cols2[c("CP","CNP","NCP","NCNP")],pch=15,xpd = TRUE,ncol=2)

```

```{r highDivQTLs8super, fig.cap="Final generation frequencies based on the model with 8 QTLs in a supergene, with parameters that are expected to result in high diversity based on the mathematical model. In 'monogamy' cases, the males were constrained to only accept eggs from one female into their nests, whereas under 'polygyny' males could mate with multiple females. In both cases females mated once. Parameter settings lablled with 'nm' were runs where females did not mate if they could not find a suitable mate (i.e., if no courters were available)."}

var_patts<-unique(gsub(".*highDiversity_(.*)_\\d_summary.txt_\\d","\\1",
                       lnames[grep(paste0("q",qs[4]),lnames)]))


par(mfrow=c(2,ceiling(length(var_patts)/2)),oma=c(3,1,4,1),mar=c(3,1,4,1),xpd=TRUE)
varDat<-lapply(var_patts,function(pattern){
  dat<-lowDiv[grep(paste0(pattern,"_\\d"),rownames(lowDiv)),]
  bp<-barplot(t(as.matrix(dat[,c("FreqNcNp", "FreqCNp", "FreqNcP", "FreqCP")])),
        col=cols2[c("NCNP","CNP","NCP","CP")],
        names.arg =gsub("^.*(\\d)_summary.txt.*$","\\1",rownames(dat)),
        las=2,
        border=NA,
        main=gsub("qtls_monogamy_nm_(.*)","\\1",pattern))
  text(x=bp,y=0.5,srt=90,labels = dat$FreqCP)

  return(dat)
})
par(fig=c(0, 1, 0, 1), oma=c(0, 0, 0, 0), mar=c(0, 0, 0, 0), new=TRUE)
plot(0,0, type='n', bty='n', xaxt='n', yaxt='n')
legend("top",bty='n',legend = c("Courter/Parent","Courter/Non-parent","Non-courter/Parent","Non-courter/Non-parent"),
       col=cols2[c("CP","CNP","NCP","NCNP")],pch=15,xpd = TRUE,ncol=2)

```

# Selection on traits

```{r traitsFiles}
traits_files<-c(
  list.files(path="../fixedART-results/qtls/", pattern="_traits.txt", full.names = TRUE),
  list.files(path="../fixedART-results/supergene/", pattern="_traits.txt", full.names = TRUE))
```

```{r traitsSelection, warning=FALSE, eval=FALSE}

seldat<-lapply(traits_files, function(filename){
  traits<-read.delim(filename)
  seldat<-do.call(rbind,by(traits, traits$Pop, function(popdat){
    
    gens<-as.matrix(do.call(rbind,by(popdat, popdat$Gen, function(dat){
      
      # selection on males
      maldat<-dat[dat$Sex=="MALE",]
      courterLM<-summary(lm(maldat$Courter ~ maldat$LifetimeRS))
      courterNumLM<-summary(lm(maldat$CourtTrait ~ maldat$LifetimeRS))
      parentLM<-summary(lm(maldat$Parent ~ maldat$LifetimeRS))
      parentNumLM<-summary(lm(maldat$ParentTrait ~ maldat$LifetimeRS))
      
      cp_freq<-nrow(maldat[maldat$Courter==1 & maldat$Parent == 1,])/nrow(maldat)
      np_freq<-nrow(maldat[maldat$Courter==0 & maldat$Parent == 1,])/nrow(maldat)
      cn_freq<-nrow(maldat[maldat$Courter==1 & maldat$Parent == 0,])/nrow(maldat)
      nn_freq<-nrow(maldat[maldat$Courter==0 & maldat$Parent == 0,])/nrow(maldat)
      
      return(c(
        court_cat_b0 = courterLM$coefficients[1,1],
        court_cat_b0_se = courterLM$coefficients[1,2],
        court_cat_b0_p = courterLM$coefficients[1,4],
        court_cat_b1 = courterLM$coefficients[2,1],
        court_cat_b1_se = courterLM$coefficients[2,2],
        court_cat_b1_p = courterLM$coefficients[2,4],
        
        parent_cat_b0 = parentLM$coefficients[1,1],
        parent_cat_b0_se = parentLM$coefficients[1,2],
        parent_cat_b0_p = parentLM$coefficients[1,4],
        parent_cat_b1 = parentLM$coefficients[2,1],
        parent_cat_b1_se = parentLM$coefficients[2,2],
        parent_cat_b1_p = parentLM$coefficients[2,4], 
        
        cp_freq = cp_freq,
        np_freq = np_freq,
        cn_freq = cn_freq,
        nn_freq = nn_freq
      ))
    })))
    
    gens<-cbind(gens, 
                gen=as.numeric(rownames(gens)),
                pop= as.numeric(popdat$Pop[1]))
    
    return(gens)
  }))
  if(length(grep("supergene", filename)>0)){
    seldat<-cbind(seldat,supergene=1)  
  } else {
    seldat<-cbind(seldat,supergene=0)
  }
  
   if(length(grep("high", filename)>0)){
    seldat<-cbind(seldat,highDiv=1)  
  } else {
    seldat<-cbind(seldat,highDiv=0)
  }
  
  return(seldat)
})
names(seldat)<-traits_files
saveRDS(seldat,"selection_gradients.RDS")

```
```{r seldat}
seldat<-readRDS("selection_gradients.RDS")
seldat<-seldat[grep("monogamy_nm",names(seldat))]
```



```{r selectionGradients, fig.height=5, fig.width=7, fig.cap="Selection gradients on parent and courter traits differ across the scenarios of our simulation model. In the first generation, selection always favours the parent trait, but selection gradients on the courter trait can be negative under parameter settings that favoured high diversity. By the end of the simulations (generation 12000), selection has been relaxed, so that selection on the parent trait is often near zero, as is selection on the courter trait. In some cases, the selection gradients on courter and parent triats are correlated at this final generation, especially in low diversity scenarios with QTLs as the genetic architecture. Solid grey lines show the x- and y-axes, with dotted grey lines showing the one-to-one correlation lines."}
par(mfrow=c(1,2))
plot(-0.25:0.25, -0.25:0.25, 
     type='n',
     xlim=c(-.25,0.25),
     ylim=c(-0.01,0.25),
     xlab="Initial generation",
     ylab="Selection gradient on parent trait",
     main="",
     bty='n', 
     axes=FALSE)
mtext("Selection gradient on courter trait", side=1,outer=TRUE, line=-3)
abline(h=0, col="grey")
abline(v=0, col="grey")
abline(a=0,b=1,col="grey",lwd=2,lty=2)
abline(a=0,b=-1,col="grey",lwd=2,lty=2)
tmp<-lapply(seldat, function(dat){
  dat<-dat[dat[,"gen"]==0,]
  
  if(dat[1,"highDiv"]==0){
    color <- "forest green"
  } else{
    color <- "cornflower blue"
  }
  points(dat[,"court_cat_b1"],
         dat[,"parent_cat_b1"],
         pch=dat[,"supergene"]+15,
         col=scales::alpha(color,0.5),
           cex=1.5
  )
  arrows(
    x0 = dat[,"court_cat_b1"] - dat[,"court_cat_b1_se"],
    x1 = dat[,"court_cat_b1"] + dat[,"court_cat_b1_se"],
    y0 = dat[,"parent_cat_b1"],
    y1 = dat[,"parent_cat_b1"],
    code = 0,
    col=scales::alpha(color,0.5)
      )
    arrows(
    x0 = dat[,"court_cat_b1"] ,
    x1 = dat[,"court_cat_b1"] ,
    y0 = dat[,"parent_cat_b1"] - dat[,"parent_cat_b1_se"],
    y1 = dat[,"parent_cat_b1"] + dat[,"parent_cat_b1_se"],
    code = 0,
    col=scales::alpha(color,0.5)
      )

})

plot(-0.25:0.25, -0.25:0.25, 
     type='n',
     xlim=c(-.25,0.25),
     ylim=c(-0.01,0.25),
     xlab="Final Generation",
     ylab="",
     main="",
     bty='n', 
     axes=FALSE)
abline(h=0, col="grey")
abline(v=0, col="grey")
abline(a=0,b=1,col="grey",lwd=2,lty=2)
abline(a=0,b=-1,col="grey",lwd=2,lty=2)
tmp<-lapply(seldat, function(dat){
  dat<-dat[dat[,"gen"]==12000,]
  if(nrow(dat)>0){
    
    if(dat[1,"highDiv"]==0){
      color <- "forest green"
    } else{
      color <- "cornflower blue"
    }
    points(dat[,"court_cat_b1"],
           dat[,"parent_cat_b1"],
           pch=dat[,"supergene"]+15,
           col=scales::alpha(color,0.5),
           cex=1.5
    )
    arrows(
      x0 = dat[,"court_cat_b1"] - dat[,"court_cat_b1_se"],
      x1 = dat[,"court_cat_b1"] + dat[,"court_cat_b1_se"],
      y0 = dat[,"parent_cat_b1"],
      y1 = dat[,"parent_cat_b1"],
      code = 0,
      col=scales::alpha(color,0.5)
        )
      arrows(
      x0 = dat[,"court_cat_b1"] ,
      x1 = dat[,"court_cat_b1"] ,
      y0 = dat[,"parent_cat_b1"] - dat[,"parent_cat_b1_se"],
      y1 = dat[,"parent_cat_b1"] + dat[,"parent_cat_b1_se"],
      code = 0,
      col=scales::alpha(color,0.5)
        )
  }
})
outer_legend("top",
       bty='n',
       pch=c(15, 16, 0,0),
       lty=c(0, 0, 1,1),
       lwd=c(2,2,1,1),
       col=c("grey", "grey","cornflower blue", "forest green"),
       legend=c("Supergenes", "QTLs", "High Diversity", "Low Diversity"),
       ncol=4)
```

What's driving these different sets of selection patterns?

```{r}
neg_sel<-unlist(lapply(seldat,function(dat){
  if(nrow(dat[dat[,"court_cat_b1"] < 0 & dat[,"gen"] == 0,])>0){
    return(TRUE)
  } else{
    return(FALSE)  
  }
}))
```


Do negative selection gradients on courters persist over time? 

```{r selectionOverTime, fig.height=4, fig.width=4, fig.cap="Selection gradients on courters do not persist over time. By generation 12000, the selection gradient is unlikely to be the same value or same direction as it was in generation 0. This pattern is true regardless of whether the parameters favoured high diversity vs low diversity or whether the genetic mechanism was QTLs or supergenes."}
plot(-0.25:0.25, -0.25:0.25, 
     type='n',
     xlim=c(-.25,0.25),
     ylim=c(-0.25,0.25),
     xlab="Selection gradient on courter trait in gen 0",
     ylab="Selection gradient on courter trait in gen 12000",
     bty='n', 
     axes=FALSE)
abline(h=0, col="grey")
abline(v=0, col="grey")
abline(a=0,b=1,col="grey",lwd=2,lty=2)
abline(a=0,b=-1,col="grey",lwd=2,lty=2)
cors<-lapply(seldat, function(dat){
  
 if(dat[1,"highDiv"]==0){
    color <- "forest green"
  } else{
    color <- "cornflower blue"
  }
  if(nrow(dat[dat[,"gen"]==12000,])>0){
     points(dat[dat[,"gen"]==0,"court_cat_b1"],
       dat[dat[,"gen"]==12000,"court_cat_b1"],
       pch=dat[,"supergene"]+15,
       col=scales::alpha(color,0.5),
       cex=1.5
  )
 
  
  arrows(
    x0 = dat[dat[,"gen"]==0,"court_cat_b1"] - dat[dat[,"gen"]==0,"court_cat_b1_se"],
    x1 = dat[dat[,"gen"]==0,"court_cat_b1"] + dat[dat[,"gen"]==0,"court_cat_b1_se"],
    y0 = dat[dat[,"gen"]==12000,"court_cat_b1"],
    y1 = dat[dat[,"gen"]==12000,"court_cat_b1"],
    code = 0,
    col=scales::alpha(color,0.5)
      )
    arrows(
    x0 = dat[dat[,"gen"]==0,"court_cat_b1"] ,
    x1 = dat[dat[,"gen"]==0,"court_cat_b1"] ,
    y0 = dat[dat[,"gen"]==12000,"court_cat_b1"] - dat[dat[,"gen"]==12000,"court_cat_b1_se"],
    y1 = dat[dat[,"gen"]==12000,"court_cat_b1"] + dat[dat[,"gen"]==12000,"court_cat_b1_se"],
    code = 0,
    col=scales::alpha(color,0.5)
      )
    
    rxy<-cor(dat[dat[,"gen"]==0,"court_cat_b1"],dat[dat[,"gen"]==12000,"court_cat_b1"])
    return(rxy)
  } else{
    return(NA)
  }
  
 
})
legend("top",
       bty='n',
       pch=c(15, 16, 0,0),
       lty=c(0, 0, 1,1),
       lwd=c(2,2,1,1),
       col=c("grey", "grey","cornflower blue", "forest green"),
       legend=c("Supergenes", "QTLs", "High Diversity", "Low Diversity"),
       ncol=2)
```

The relationship between selection gradients at generation 0 and generation 12000 follow a roughly uniform distribution (Fig. \@ref(fig:gensSelectionCorHist)) -- there is no pattern.

```{r gensSelectionCorHist, fig.height=4, fig.width=4, fig.cap="Distribution of correlation coefficients between generation 0 and generation 12000 selection gradients on the courter trait."}
hist(unlist(cors),
     xlab="Correlation coefficient", 
     ylab="Number of replicates",
     main="")
```

Does selection at generation 0 predict diversity at generation 12000?

```{r selFreq, fig.height=6, fig.width=6, fig.cap="The direction of selection on the courter trait in generation zero predicts whether diversity of morphs will be maintained."}
plot(
  seq(-0.15,0.15,length.out=2),
  0:1,
  type='n',
  bty='l',
  xlab="Selection gradient on courters in generation 0",
  ylab="Final Courter/Parent frequency")
divSel<-lapply(seldat, function(dat){
  # put the two gens next to each other
  mixdat<-cbind(
    selgrad=dat[dat[,"gen"]==0,"court_cat_b1"],
    cpfreq=dat[dat[,"gen"]==12000,"cp_freq"],
    highDiv=dat[dat[,"gen"]==0,"highDiv"],
    supergene=dat[dat[,"gen"]==0,"supergene"]
    )
  
  
  if(mixdat[1,"highDiv"]==0){
    color <- "forest green"
  } else{
    color <- "cornflower blue"
  }
  points(mixdat[,1],
     mixdat[,2],
     pch=mixdat[,"supergene"]+15,
     col=scales::alpha(color,0.5),
     cex=1.5,
     xpd=TRUE
  )
  
  return(mixdat)
})
legend("topleft",
       bty='n',
       pch=c(15, 16, 0,0),
       lty=c(0, 0, 1,1),
       lwd=c(2,2,1,1),
       col=c("grey", "grey","cornflower blue", "forest green"),
       legend=c("Supergenes", "QTLs", "High Diversity", "Low Diversity"),
       ncol=2)
```

Which of these ones have cp_freq close to 1 and 0?? They haven't shown up in the plots of high diversity parameters.

```{r}
for(i in 1:length(seldat)){
  if(mean(seldat[[i]][seldat[[i]][,"gen"]==12000,"cp_freq"])>0.9){
    browser()
    dat<-get.morph.freqs(
      plot.pc.reps(pattern=gsub("../fixedART-results/.*//(.*)_traits.txt","\\1_summary.txt",traits_files[i]),
                   path=gsub("(../fixedART-results/.*/).*$","\\1",traits_files[i]),
                   cols=cols,make.plot=FALSE))
    
    bp<-barplot(t(as.matrix(dat[,c("FreqNcNp", "FreqCNp", "FreqNcP", "FreqCP")])),
        col=cols2[c("NCNP","CNP","NCP","CP")],
        names.arg =gsub("^.*(\\d)_summary.txt.*$","\\1",rownames(dat)),
        las=2,
        border=NA,
        main=traits_files[i])
  text(x=bp,y=0.5,srt=90,labels = dat$FreqCP)

    
  } 
}

```


### Evaluating balancing selection

While we can look at the selection gradients on the traits in general, we might also want to know if the surviving morphs have similar reproductive success. 


fitness_dat<-lapply(traits_files, function(filename){
  traits<-read.delim(filename)
  fitdat<-do.call(rbind,by(traits, traits$Pop, function(popdat){
    
    gens<-as.matrix(do.call(rbind,by(popdat, popdat$Gen, function(dat){
      # males
      maldat<-dat[dat$Sex=="MALE",]
      
      return(c(
        cp_muRS=mean(maldat$LifetimeRS[maldat$Courter==1 & maldat$Parent == 1]),
        np_muRS=mean(maldat$LifetimeRS[maldat$Courter==0 & maldat$Parent == 1]),
        cn_muRS=mean(maldat$LifetimeRS[maldat$Courter==1 & maldat$Parent == 0]),
        nn_muRS=mean(maldat$LifetimeRS[maldat$Courter==0 & maldat$Parent == 0]),
        
        cp_semRS=sem(maldat$LifetimeRS[maldat$Courter==1 & maldat$Parent == 1]),
        np_semRS=sem(maldat$LifetimeRS[maldat$Courter==0 & maldat$Parent == 1]),
        cn_semRS=sem(maldat$LifetimeRS[maldat$Courter==1 & maldat$Parent == 0]),
        nn_semRS=sem(maldat$LifetimeRS[maldat$Courter==0 & maldat$Parent == 0]),
        
        cp_freq=nrow(maldat[maldat$Courter==1 & maldat$Parent == 1,])/nrow(maldat),
        np_freq=nrow(maldat[maldat$Courter==0 & maldat$Parent == 1,])/nrow(maldat),
        cn_freq=nrow(maldat[maldat$Courter==1 & maldat$Parent == 0,])/nrow(maldat),
        nn_freq=nrow(maldat[maldat$Courter==0 & maldat$Parent == 0,])/nrow(maldat)
        
      ))
    })))
     
    gens<-cbind(gens, 
                gen=as.numeric(rownames(gens)),
                pop= as.numeric(popdat$Pop[1]))
    
    return(gens)
  }))
  if(length(grep("supergene", filename)>0)){
    fitdat<-cbind(fitdat,supergene=1)  
  } else {
    fitdat<-cbind(fitdat,supergene=0)
  }
  
   if(length(grep("high", filename)>0)){
    fitdat<-cbind(fitdat,highDiv=1)  
  } else {
    fitdat<-cbind(fitdat,highDiv=0)
  }
  
  return(fitdat)
})
names(fitness_dat)<-traits_files

```

```{r}

plot(1:4,
     seq(0,3,length.out=4),
     xlab="",
     xaxt='n',
     ylab="Lifetime Reproductive Success",
     bty='l',
     type='n')
axis(1,at=1:4,labels=c("CP","NP","CN","NN"))

for(i in 1:length(fitness_dat)){
   
  fitdat12000<-fitness_dat[[i]][fitness_dat[[i]][,"gen"]==12000,]
  
  if(nrow(fitdat12000)>0){
    
    apply(fitdat12000,1,function(dat){
      # points
      points(1:4,
         dat[1:4],
         pch=19,
         col=scales::alpha(cols2,0.5))
      arrows(x0=1:4,
           x1=1:4,
           y0=dat[1:4]+dat[5:8],
           y1=dat[1:4]-dat[5:8],
           code=0,
           col=scales::alpha(cols2,0.5)
           )
      lines(1:4,
            dat[1:4],
            lwd=2,
            col=scales::alpha("grey",0.5))
      })
    
 
    }
}



```


