---
title: "Visualising results using ternary diagrams to compare all model outcomes"
output:
  pdf_document: default
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,out.extra='',fig.pos="H",dpi=200,fig.height = 7,fig.width = 7)
knitr::opts_knit$set(root.dir='../results/') #change
```
```{r opts, echo = FALSE}
library(knitr)
knitr::opts_chunk$set(
  fig.path = "../figs/",
  dev="png"
)
```
```{r librarysetup, echo=FALSE,echo=FALSE,message=FALSE}
library(RColorBrewer)
library(scales)
library(sm)
library(kableExtra)
library(plotly)
library(dplyr)
library(vegan)
library(tidyr)
library(Ternary)
source("../R/002_freq_functions.R")
source("../morph_predictions/check_freqs.R")
source("../morph_predictions/morph_gens_ns.R")
source("../R/formatting.R")
```

```{r spftools}
if(length(system.file(package='spfTools'))<=1){
  devtools::install_github("https://github.com/spflanagan/spfTools/")
}
library(spfTools)
```


```{r dataConversionFXN}
morphs_to_ternary<-function(pattern, path){
  data<-get.morph.freqs(plot.pc.reps(pattern=pattern,path=path,cols,make.plot=FALSE))
  data_points<-apply(data,
                   1,
                   function(dat){
                     out<-dat[c("FreqCP","FreqNcP","FreqNcNp")]
                     return(as.numeric(out))
                   },
                   simplify = FALSE)
  return(data_points)
}

```
```{r makeTernaryMorphsFXN}
# made with inspiration from this tutorial: https://cran.r-project.org/web/packages/Ternary/vignettes/Ternary.html
morphs_ternary_plot<-function(...){
  TernaryPlot(atip="",
            btip="",
            ctip="",
            alab="Courter/Parent",
            blab="Noncourter/Parent",
            clab="Noncourter/Nonparent",
            grid.lines=4,
            grid.minor.lines = 1,
            grid.lty = "dotted",
            grid.minor.lty = "dotted",
            lab.col=cols2[c(1,2,4)],
            tip.col=cols2[c(1,2,4)],
            ...)
}
```

One of the challenges with the model outputs are that they rely on frequencies that all relate to each other. Therefore, we will use ternary diagrams to more easily see patterns and compare final frequencies. 

## Baseline model

The baseline model results are saved in `morph_results_Ns.RDS`, so we can rapidly and easily load those into R.

```{r baselineData}
morph_results<-readRDS("../morph_predictions/morph_results_Ns.RDS")
morph_results$diversity<-vegan::diversity(round(
  morph_results[,c("CP","CN","NP","NN")],4))

baseline_points<- morph_results[which(morph_results$initial_CP==0.25 &
                                        morph_results$initial_NP==0.25 & 
                                        morph_results$initial_NN==0.25),
                                c("CP","NP","NN")]
baseline_points<-baseline_points[which(rowSums(baseline_points)>0),]

```

```{r baselineTernary, eval=FALSE}
morphs_ternary_plot()

# add the points
AddToTernary(
  points,
  baseline_points,
  col=alpha("grey",0.5),
  lwd=2,
  cex=2 # add color based on param values
)
```



```{r dupdat}
dupdat<-data.table::setDT(baseline_points)[,list(Count=.N),names(baseline_points)]
# Calculate the percentage for easier point plotting
dupdat$Percent<-dupdat$Count/sum(dupdat$Count)
```


```{r ternaryScaledPts, eval=FALSE}
# Make the ternary diagram
morphs_ternary_plot()

# add the points
AddToTernary(
  points,
  col="purple",
  dupdat[,1:3],
  lwd=2,
  cex=5*dupdat$Percent
)
```


## Single locus simulations

These simulations had three diversity settings but also four mating systems, so I'll use a mix of colors and points to show the differences.

```{r singleLocusData, eval=FALSE}
lowDiv<-morphs_to_ternary(pattern="lowDiversity.*summary.txt",path="single_locus")
highDiv<-morphs_to_ternary(pattern="highDiversity_.*summary.txt", path="single_locus")
highDivStrict<-morphs_to_ternary(pattern="highDiversityStrict.*summary.txt",path="single_locus")

# remove ones with no viability sel or random mating
lowDiv<-lowDiv[grep("v0",grep("RM", names(lowDiv), invert=TRUE, value=TRUE), invert=TRUE, value=TRUE)] 
highDiv<-highDiv[grep("v0",grep("RM", names(highDiv), invert=TRUE, value=TRUE), invert=TRUE, value=TRUE)]
highDivStrict<-highDivStrict[grep("v0",grep("RM", names(highDivStrict), invert=TRUE, value=TRUE), invert=TRUE, value=TRUE)]
saveRDS(lowDiv,"lowDiv.RDS")
saveRDS(highDiv,"highDiv.RDS")
saveRDS(highDivStrict,"highDivStrict.RDS")
```
```{r readSingleLocusData}
lowDiv<-readRDS("lowDiv.RDS")
highDiv<-readRDS("highDiv.RDS")
highDivStrict<-readRDS("highDivStrict.RDS")
```


```{r singleLocusMSinfo}
# extract corresponding mating system info
low_patts<-gsub(".*lowDiversity_(.*)_\\d_summary.txt_\\d","\\1",names(lowDiv))
low_patts<-grep("v0",grep("RM", low_patts, invert=TRUE, value=TRUE), invert=TRUE, value=TRUE)

high_patts<-gsub(".*highDiversity_(.*)_\\d_summary.txt_\\d","\\1",names(highDiv))
high_patts<-grep("v0",grep("RM", high_patts, invert=TRUE, value=TRUE), invert=TRUE, value=TRUE)

strict_patts<-gsub(".*highDiversityStrict_(.*)_\\d_summary.txt_\\d","\\1",names(highDivStrict))
strict_patts<-grep("v0",grep("RM", strict_patts, invert=TRUE, value=TRUE), invert=TRUE, value=TRUE)

```


```{r singleLocusPlot, eval=FALSE}


# create the base
par(mar=c(0,0,0,0))
morphs_ternary_plot()

# add the points
AddToTernary(
  points,
  lowDiv,
  col=ms_cols[low_patts],
  pch=6,
  lwd=2,
  cex=2 # add color based on param values
)

AddToTernary(
  points,
  highDiv,
  col=ms_cols[high_patts],
  pch=0,
  lwd=2,
  cex=2 # add color based on param values
)

AddToTernary(
  points,
  highDivStrict,
  col=ms_cols[strict_patts],
  pch=5,
  lwd=2,
  cex=2 # add color based on param values
)

outer_legend("topleft",
       c(names(ms_cols),"low", "high", "strict high"),
       col=c(ms_cols, rep("dark grey",3)),
       pch=c(rep(1,4),6,0,5),
       pt.cex=2,
       bty='n',
       ncol=2,
       pt.lwd=2)


```


```{r singleLocusSidBySide, eval=FALSE}


# create the base
par(mar=c(0,0,0,0),
    mfrow=c(1,4))
morphs_ternary_plot()

# add the points
AddToTernary(
  points,
  lowDiv,
  col=ms_cols[low_patts],
  pch=6,
  lwd=2,
  cex=2 # add color based on param values
)

morphs_ternary_plot()
AddToTernary(
  points,
  highDiv,
  col=ms_cols[high_patts],
  pch=0,
  lwd=2,
  cex=2 # add color based on param values
)

morphs_ternary_plot()
AddToTernary(
  points,
  highDivStrict,
  col=ms_cols[strict_patts],
  pch=5,
  lwd=2,
  cex=2 # add color based on param values
)

outer_legend("topleft",
       c(names(ms_cols),"low", "high", "strict high"),
       col=c(ms_cols, rep("dark grey",3)),
       pch=c(rep(1,4),6,0,5),
       pt.cex=2,
       bty='n',
       ncol=2,
       pt.lwd=2)


```

```{r monogamyPlot, eval=FALSE}


# create the base
par(mar=c(0,0,0,0))
morphs_ternary_plot()

# add the points
AddToTernary(
  points,
  lowDiv,
  col=ms_cols[low_patts],
  pch=6,
  lwd=2,
  cex=2 # add color based on param values
)

AddToTernary(
  points,
  highDiv,
  col=ms_cols[high_patts],
  pch=0,
  lwd=2,
  cex=2 # add color based on param values
)

AddToTernary(
  points,
  highDivStrict,
  col=ms_cols[strict_patts],
  pch=5,
  lwd=2,
  cex=2 # add color based on param values
)

outer_legend("topleft",
       c(names(ms_cols),"low", "high", "strict high"),
       col=c(ms_cols, rep("dark grey",3)),
       pch=c(rep(1,4),6,0,5),
       pt.cex=2,
       bty='n',
       ncol=2,
       pt.lwd=2)


```


I want to split out the low and high but also the four mating types. 

```{r}
lowReps<-gsub("^.*_(\\d)_summary.txt.*$","\\1",names(lowDiv))
highReps<-gsub("^.*_(\\d)_summary.txt.*$","\\1",names(highDiv))

labels<-c(
  monogamy="monogamy \n+ random",
  monogamy_nm="monogamy\n          ",
  polygyny="polygyny \n+ random",
  polygyny_nm="polygyny\n         "
)
```


```{r matingSystemsTernary, fig.height=5.5, fig.width=7}
# create the base
par(mar=c(0,2,0,1.5), 
    oma=c(1.5,2.5,3,0),
    mfrow=c(2,4), 
    xpd=TRUE)

# Low diversity plots
for(ms in unique(low_patts)){
  
  # identify the ones
  indx<-which(low_patts %in% ms)
  
  # make the base plot
  morphs_ternary_plot(lab.cex=1.5,
                      lab.offset=0.25,
                      atip.pos=3,
                      atip.rotate=0,
                      btip.pos=2, 
                      btip.rotate=0,
                      ctip.pos=2,
                      ctip.rotate=0)

  # add the points
  AddToTernary(
    points,
    lowDiv[indx],
    col=ms_cols[low_patts[indx]],
    pch=as.numeric(lowReps[indx]),
    lwd=2,
    cex=2 # add color based on param values
  )
  mtext(labels[ms],3, line = 2)
}

par(mar=c(1,2,0,1.5))
# high diversity plots
for(ms in unique(high_patts)){
  
  # identify the ones
  indx<-which(high_patts %in% ms)
  
  # make the base plot
  morphs_ternary_plot(lab.cex=1.5,
                      lab.offset=0.25,
                      atip.pos=3,
                      atip.rotate=0,
                      btip.pos=2, 
                      btip.rotate=0,
                      ctip.pos=2,
                      ctip.rotate=0)

  # add the points
  AddToTernary(
    points,
    highDiv[indx],
    col=ms_cols[high_patts[indx]],
    pch=as.numeric(highReps[indx]),
    lwd=2,
    cex=2 # add color based on param values
  )
}

mtext("Low diversity", 
      side = 2, 
      outer=TRUE,
      adj=0.85, 
      col=divcols["low"],
      padj=-0.5)
mtext("High diversity", 
      side = 2,
      outer=TRUE,
      adj=0.25,
      padj=-0.5,
      col=divcols["high"])

```

```{r SingleLocusReplicates, fig.width=9, fig.height=3.5}
# narrow down to the ones we actually analysed
focus<-highDiv[grep("monogamy_nm",names(highDiv))]
reps<-gsub("^.*_(\\d)_summary.txt.*$","\\1",names(focus))


par(mar=c(1,2,0,1.5), 
    oma=c(1,0,1,0),
    mfrow=c(1,length(unique(reps))), 
    xpd=TRUE)

# make the plots
for(rep in unique(reps)){
  
  # identify the ones
  indx<-grep(paste0(rep,"_summary"),names(focus))
  
  
  # make the base plot
  morphs_ternary_plot(lab.cex=1.5,
                      lab.offset=0.25,
                      atip.pos=3,
                      atip.rotate=0,
                      btip.pos=2, 
                      btip.rotate=0,
                      ctip.pos=2,
                      ctip.rotate=0)

  # add the points
  AddToTernary(
    points,
    focus[indx],
    col=ms_cols["monogamy_nm"],
    pch=as.numeric(rep),
    lwd=2,
    cex=2 # add color based on param values
  )
  
}

```


## Polygenic inheritance

The main variables I want to show with the polygenic inheritance are the genetic architecture (QTLs vs supergenes with diff props) and number of chromosomes. I could also do number of QTLs -- that wasn't shown in the main manuscript, as I showed representative images from the 8 QTLs scenarios. How do I want to show this? 

- In other figures, I've used circles and triangles for QTLs and supergenes, respectively
- I could use colors to show num of chromosomes and do different triangles for each architecture
- scale points based on the number of QTLs?

```{r finalfreqs}
final_freqs<-read.csv("../results/morph_freqs_summary.csv")
final_freqs$rep<-gsub("^.*Diversity_(.*)_pheno.csv","\\1",final_freqs$file)

# extract the relevant info for plotting
final_freqs$nqtl<-as.numeric(gsub("^.*q(\\d+)_.*$","\\1",final_freqs$rep))
final_freqs$nchrom<-as.factor(gsub("^.*c(\\d+)_.*$","\\1",final_freqs$rep))
final_freqs$arch<-gsub("^.*results\\/(\\w+)\\/.*$","\\1",final_freqs$file)
final_freqs$arch[final_freqs$arch=="supergene"]<-gsub("^.*(prop\\d+\\.\\d+).*$","\\1",final_freqs$file[final_freqs$arch=="supergene"])
final_freqs$diversity<-gsub("^.*\\/(\\w+)Diversity.*$","\\1",final_freqs$file)

# only keep the high diversity ones
high_freqs<-final_freqs[which(final_freqs$diversity=="high"),]

```



I'll start with just the QTL options.

```{r getpoints}

points_list<-by(high_freqs, high_freqs$arch, function(arch_dat){
  # convert the data into the list format
  points<-apply(arch_dat,
               1,
               function(dat){
                 out<-dat[c("CP","P","NON")]
                 return(as.numeric(out))
                 },
               simplify = FALSE)
  # output the data with the other info
  return(list(points=points,
              chroms=as.factor(arch_dat$nchrom),
              qtls=as.numeric(arch_dat$nqtl),
              arch=unique(arch_dat$arch)
              )
         )
}
)




```


```{r architectureTernary}
arch_pch<-c(rep(2,3),6)
names(arch_pch)<-names(points_list)


par(mar=c(0,0,0,0),
    mfrow=c(1,4))

tmp<-lapply(points_list, function(points_info){
  morphs_ternary_plot()
  
  # add the points
  AddToTernary(
    points,
    points_info$points,
    col=alpha(chr_col[points_info$chroms],0.75),
    pch=arch_pch[points_info$arch],
    lwd=2,
    cex=log(points_info$qtls,base = 2)-1
  )
  text(x=0,
       y=1.1,
       points_info$arch,
       cex=2,
       xpd=TRUE
       )
})



lgd<-outer_legend(x=-0.35,
                  y=0.65,
                  names(chr_col),
                  col=chr_col,
                  pch=15,
                  cex=2,
                  bty='n',
                  horiz=TRUE,
                  pt.lwd=2,
                  title="Number of Chromosomes")
outer_legend(x=-0.35,
             y=-0.45,
             legend=levels(as.factor(unique(high_freqs$nqtl))),
             pch=15,
             cex=2,
             pt.cex=log(as.numeric((levels(as.factor(unique(high_freqs$nqtl))))),base=2)-1,
             horiz=2,
             pt.lwd=2,
             bty='n',
             title="Number of QTLs")
```





## A combined diagram

I will create a combined diagram, with the top row showing the baseline model, the single locus model, and QTLs, and the second row showing the three supergene options. Maybe add a star to the baseline model to show which parameter combination we selected to use for the simulations?

I'll only want to plot the monogamy_nm ones from the single_locus runs.

```{r onlyMonogNM}
sim_high<-morph_results[morph_results$initial_CN==0.25 & 
                           morph_results$initial_CP==0.25 &
                           morph_results$initial_NN==0.25 &
                           morph_results$initial_NP==0.25 &
                           round(morph_results$c,2) == 0.75 &
                           round(morph_results$r,1) == 0.7 &
                           morph_results$num_sneak == 2,]

sim_low<-morph_results[morph_results$initial_CN==0.25 & 
                           morph_results$initial_CP==0.25 &
                           morph_results$initial_NN==0.25 &
                           morph_results$initial_NP==0.25 &
                           round(morph_results$c,2) == 0.5 &
                           round(morph_results$r,1) == 2 &
                           morph_results$num_sneak == 2,]
```


```{r combinedTernary, include=FALSE}

par(mar=c(0,0,0,0),
    mfrow=c(2,3))


# --------------- Baseline model
dupdat<-data.table::setDT(baseline_points)[,list(Count=.N),names(baseline_points)]
# Calculate the percentage for easier point plotting
dupdat$Percent<-dupdat$Count/sum(dupdat$Count)

# Make the ternary diagram
morphs_ternary_plot()

# add the points
AddToTernary(
  points,
  col="black",
  dupdat[,1:3],
  lwd=2,
  cex=6*dupdat$Percent
)


# -------------- Single locus sims

morphs_ternary_plot()

# add the points
AddToTernary(
  points,
  lowDiv[low_patts=="monogamy_nm"],
  col="black",
  pch=6,
  lwd=2,
  cex=2 # add color based on param values
)

AddToTernary(
  points,
  highDiv[low_patts=="monogamy_nm"],
  col="black",
  pch=0,
  lwd=2,
  cex=2 # add color based on param values
)

AddToTernary(
  points,
  highDivStrict[strict_patts=="monogamy_nm"],
  col="black",
  pch=5,
  lwd=2,
  cex=2 # add color based on param values
)

# outer_legend("top",
#        c("low", "high", "strict high"),
#        col="black",
#        pch=c(6,0,5),
#        pt.cex=2,
#        bty='n',
#        ncol=3,
#        pt.lwd=2)


# ---------------- QTLs etc. 

tmp<-lapply(points_list[c("qtls","prop0.05","prop0.25","prop0.5")], 
            function(points_info){
              morphs_ternary_plot()
              
              # add the points
              AddToTernary(
                points,
                points_info$points,
                col=alpha(chr_col[points_info$chroms],0.75),
                pch=arch_pch[points_info$arch],
                lwd=2,
                cex=log(points_info$qtls,base = 2)-1
              )
              text(x=0,
                   y=1.1,
                   points_info$arch,
                   cex=2,
                   xpd=TRUE
                   )
})



# lgd<-outer_legend(x=-0.35,
#                   y=0.65,
#                   names(chr_col),
#                   col=chr_col,
#                   pch=15,
#                   cex=2,
#                   bty='n',
#                   horiz=TRUE,
#                   pt.lwd=2,
#                   title="Number of Chromosomes")
# outer_legend(x=-0.35,
#              y=-0.45,
#              legend=levels(as.factor(unique(high_freqs$nqtl))),
#              pch=15,
#              cex=2,
#              pt.cex=log(as.numeric((levels(as.factor(unique(high_freqs$nqtl))))),base=2)-1,
#              horiz=2,
#              pt.lwd=2,
#              bty='n',
#              title="Number of QTLs")

```


I'm not sure I like the fully combined one, as it hides some of the important details (low vs high diversity settings, for example.) What if instead I have two plots, one with the baseline and low/high/strict settings? 

```{r separateDataTypes}
base_dat<-morph_results[which(morph_results$initial_CP==0.25 &
                                        morph_results$initial_NP==0.25 & 
                                        morph_results$initial_NN==0.25),
                                ]
high_which<-which(base_dat$initial_CN==0.25 & 
                           base_dat$initial_CP==0.25 &
                           base_dat$initial_NN==0.25 &
                           base_dat$initial_NP==0.25 &
                           round(base_dat$c,2) == 0.75 &
                           round(base_dat$r,1) == 0.7 &
                           base_dat$num_sneak == 2)
low_which<-which(base_dat$initial_CN==0.25 & 
                           base_dat$initial_CP==0.25 &
                           base_dat$initial_NN==0.25 &
                           base_dat$initial_NP==0.25 &
                           round(base_dat$c,2) == 0.5 &
                           round(base_dat$r,1) == 2 &
                           base_dat$num_sneak == 2)
```


```{r baselinesWithStrict}

layout(matrix(c(1,1,2,3), nrow=2))

# --------------- Baseline model
dupdat<-data.table::setDT(baseline_points)[,list(Count=.N),names(baseline_points)]
# Calculate the percentage for easier point plotting
dupdat$Percent<-dupdat$Count/sum(dupdat$Count)

# Make the ternary diagram
morphs_ternary_plot()

# add the points
AddToTernary(
  points,
  col="black",
  dupdat[,1:3],
  lwd=2,
  cex=10*dupdat$Percent
)

#high div point
AddToTernary(
  points,
  col=divcols[1],
  baseline_points[high_which,],
  lwd=2,
  pch=8,
  cex=2
)
#low dive point
AddToTernary(
  points,
  col=divcols[2],
  baseline_points[low_which,],
  lwd=2,
  pch=8,
  cex=1
)

# -------------- Single locus sims: low

morphs_ternary_plot()

# add the points
AddToTernary(
  points,
  lowDiv[low_patts=="monogamy_nm"],
  col=divcols[2],
  pch=6,
  lwd=2,
  cex=2 # add color based on param values
)

# -------------- Single locus sims: high

morphs_ternary_plot()

AddToTernary(
  points,
  highDiv[high_patts=="monogamy_nm"],
  col=divcols[1],
  pch=0,
  lwd=2,
  cex=2 # add color based on param values
)

AddToTernary(
  points,
  highDivStrict[strict_patts=="monogamy_nm"],
  col=divcols[1],
  pch=5,
  lwd=2,
  cex=2 # add color based on param values
)
```




```{r baselinesNoStrict}

layout(matrix(c(1,1,2,3), nrow=2))

# --------------- Baseline model
dupdat<-data.table::setDT(baseline_points)[,list(Count=.N),names(baseline_points)]
# Calculate the percentage for easier point plotting
dupdat$Percent<-dupdat$Count/sum(dupdat$Count)

# Make the ternary diagram
morphs_ternary_plot()

# add the points
AddToTernary(
  points,
  col="black",
  dupdat[,1:3],
  lwd=2,
  cex=100*dupdat$Percent
)

#high div point
AddToTernary(
  points,
  col=divcols[1],
  baseline_points[high_which,],
  lwd=2,
  pch=8,
  cex=2
)
#low dive point
AddToTernary(
  points,
  col=divcols[2],
  baseline_points[low_which,],
  lwd=2,
  pch=8,
  cex=1
)

# -------------- Single locus sims: low

morphs_ternary_plot()

# add the points
AddToTernary(
  points,
  lowDiv[low_patts=="monogamy_nm"],
  col=divcols[2],
  pch=6,
  lwd=2,
  cex=2 # add color based on param values
)

# -------------- Single locus sims: high

morphs_ternary_plot()

AddToTernary(
  points,
  highDiv[high_patts=="monogamy_nm"],
  col=divcols[1],
  pch=0,
  lwd=2,
  cex=2 # add color based on param values
)


```

### The figures to use in the main text

```{r baselinesNoStrictPercentages, fig.width=8.15, fig.height=4.72}

layout(matrix(c(1,1,2,3), nrow=2))
par(mar=c(1,1.75,0.5,1), oma=c(1,3,5,10),xpd=TRUE)
# --------------- Baseline model
dupdat<-data.table::setDT(baseline_points)[,list(Count=.N),names(baseline_points)]
# Calculate the percentage for easier point plotting
dupdat$Percent<-dupdat$Count/sum(dupdat$Count)

# Make the ternary diagram
morphs_ternary_plot(lab.cex=1.5,
                      lab.offset=0.25,
                      atip.pos=3,
                      atip.rotate=0,
                      btip.pos=1, 
                      btip.rotate=0,
                      ctip.pos=2,
                      ctip.rotate=0)

# add the points
AddToTernary(
  points,
  col="black",
  dupdat[,1:3],
  lwd=2,
  cex=10*dupdat$Percent+1
)

#high div point
AddToTernary(
  points,
  col=divcols["high"],
  baseline_points[high_which,],
  lwd=3,
  pch=8,
  cex=3
)
#low div point
AddToTernary(
  points,
  col=divcols["low"],
  baseline_points[low_which,],
  lwd=3,
  pch=8,
  cex=3
)

mtext("analytical",3,cex=1.5)

# -------------- Single locus sims: low
par(mar=c(1,1.75,0.5,3))
lowdat<-as.data.frame(do.call(rbind,lowDiv[low_patts=="monogamy_nm"]))
colnames(lowdat)<-c("CP","NP","NN")
dupdat<-data.table::setDT(lowdat)[,list(Count=.N),
                                  names(lowdat)]
# Calculate the percentage for easier point plotting
dupdat$Percent<-dupdat$Count/sum(dupdat$Count)

morphs_ternary_plot(lab.cex=1.5,
                      lab.offset=0.25,
                      atip.pos=3,
                      atip.rotate=0,
                      btip.pos=1, 
                      btip.rotate=0,
                      ctip.pos=2,
                      ctip.rotate=0)

# add the points
AddToTernary(
  points,
  dupdat[,1:3],
  col=divcols["low"],
  pch=1,
  lwd=2,
  cex=10*dupdat$Percent
)

mtext("single-locus simulations",3, cex=1.5, line=0.5)

# -------------- Single locus sims: high
highdat<-as.data.frame(do.call(rbind,highDiv[high_patts=="monogamy_nm"]))
colnames(highdat)<-c("CP","NP","NN")
dupdat<-data.table::setDT(highdat)[,list(Count=.N),
                                  names(highdat)]
dupdat$Percent<-dupdat$Count/sum(dupdat$Count)

morphs_ternary_plot(lab.cex=1.5,
                      lab.offset=0.25,
                      atip.pos=3,
                      atip.rotate=0,
                      btip.pos=1, 
                      btip.rotate=0,
                      ctip.pos=2,
                      ctip.rotate=0)

AddToTernary(
  points,
  dupdat[,1:3],
  col=divcols["high"],
  pch=1,
  lwd=2,
  cex=50*dupdat$Percent # add color based on param values
)


outer_legend("topleft","(A) no explicit genetics", bty='n', cex=2, 
       x.intersp = -0.9)

outer_legend("right", 
       c("baseline","low diversity", "high diversity",
         "",
         "point size","shows freq.","of occurrence"
         ),
       pch=1,
       col=c("black",divcols[c("low","high")],NA,"dark grey","dark grey","dark grey"),
       bty='n',
       pt.cex=c(1.5,1.5,1.5,1.5,1,1.5,3),
       cex=1.5,
       pt.lwd=2
      )

# 
# # add lines to low diversity
# lines(c(-0.55,0.45),
#       c(0.6,1.02),
#       xpd=TRUE,
#       col=divcols[2],
#       lwd=2
#       )
# 
# lines(c(-0.55,0.3),
#       c(0.6,0),
#       xpd=TRUE,
#       col=divcols[2],
#       lwd=2
#       )
# 
# # add lines to high diversity
# lines(c(-0.55,0.4),
#       c(-0.12,-1.2),
#       xpd=TRUE,
#       col=divcols[1],
#       lwd=2
#       )
# 
# lines(c(-0.55,0.4),
#       c(-0.12,-0.1),
#       xpd=TRUE,
#       col=divcols[1],
#       lwd=2
#       )

```

And then to follow up on this one, we'll do a two-row genetic architectures one




```{r architectureTernaryLowHigh, fig.width=8.15, fig.height=4.72}

  arch_pch<-rep(2,4)
names(arch_pch)<-names(points_list)

arch_labels<-c(qtls="QTLs",
              prop0.05="Supergene 5%",
              prop0.25="Supergene 25%",
              prop0.5="Supergene 50%")

layout(matrix(c(1,2,3,4,5,6,7,8),nrow=2,byrow = TRUE))
par(mar=c(0,2,1,1.5),
    oma=c(1,5,3,0),
    xpd=TRUE)

# -------------- low diversity


low_freqs<-final_freqs[which(final_freqs$diversity=="low"),]
points_list_low<-by(low_freqs, low_freqs$arch, function(arch_dat){
  # convert the data into the list format
  points<-apply(arch_dat,
               1,
               function(dat){
                 out<-dat[c("CP","P","NON")]
                 return(as.numeric(out))
                 },
               simplify = FALSE)
  # output the data with the other info
  return(list(points=points,
              chroms=as.factor(arch_dat$nchrom),
              qtls=as.numeric(arch_dat$nqtl),
              arch=unique(arch_dat$arch)
              )
         )
}
)
tmp<-lapply(points_list_low[c("qtls","prop0.05","prop0.25","prop0.5")], function(points_info){
  morphs_ternary_plot(lab.cex=1.5,
                      lab.offset=0.25,
                      atip.pos=3,
                      atip.rotate=0,
                      btip.pos=2, 
                      btip.rotate=0,
                      ctip.pos=2,
                      ctip.rotate=0)
  
  # add the points
  AddToTernary(
    points,
    points_info$points,
    col=alpha(chr_col[points_info$chroms],0.75),
    pch=arch_pch[points_info$arch],
    lwd=2,
    cex=log(points_info$qtls,base = 2)-1
  )
  text(x=0,
       y=1.2,
       arch_labels[points_info$arch],
       cex=2,
       xpd=TRUE
       )
})
mtext("Low diversity",
      2,
      outer=TRUE,
      adj = 0.895,
      col=divcols["low"],
      cex=1.5)

# --------------- high diversity
par(mar=c(2,2,0,1.5))

tmp<-lapply(points_list[c("qtls","prop0.05","prop0.25","prop0.5")], function(points_info){
  morphs_ternary_plot(lab.cex=1.5,
                      lab.offset=0.25,
                      atip.pos=3,
                      atip.rotate=0,
                      btip.pos=2, 
                      btip.rotate=0,
                      ctip.pos=2,
                      ctip.rotate=0)
  
  # add the points
  AddToTernary(
    points,
    points_info$points,
    col=alpha(chr_col[points_info$chroms],0.75),
    pch=arch_pch[points_info$arch],
    lwd=2,
    cex=log(points_info$qtls,base = 2)-1
  )
})

mtext("High diversity",
      2,
      outer=TRUE,
      adj = 0.235,
      col=divcols["high"],
      cex=1.5)

lgd<-outer_legend(x=-1.15,
                  y=-0.87,
                  c("# Chrom:", names(chr_col)),
                  col=c(NA,chr_col),
                  pch=c(NA,rep(15,3)),
                  cex=2,
                  bty='n',
                  horiz=TRUE,
                  pt.lwd=2,
                 text.width=c(0.225,rep(0.1,3)))
outer_legend(x=0,
             y=-0.87,
             legend=c("# QTLs:", levels(as.factor(unique(high_freqs$nqtl)))),
             pch=c(NA,rep(2,4)),
             col=c(NA,rep("black",4)),
             cex=2,
             pt.cex=c(0,log(as.numeric((levels(as.factor(unique(high_freqs$nqtl))))),base=2)-1),
             horiz=2,
             pt.lwd=2,
             bty='n',
             text.width=c(0.2, rep(0.1,4)),
             inset=0)

outer_legend("topleft","(B) explicit genetic architecture", bty='n', cex=2.5, x.intersp = -0.9)


```


### All ternary plots together

```{r combinedTernaryTall}
figa <-grid::rasterGrob(png::readPNG("../figs/baselinesNoStrictPercentages-1.png"))
figb<-grid::rasterGrob(png::readPNG("../figs/architectureTernaryLowHigh-1.png"))
gridExtra::grid.arrange(figa, figb, nrow=2)


```



\newpage
\blandscape

```{r combinedTernaryWide, fig.width=12.225, fig.height=4.72}

layout(matrix(c(1:6, 1, 7:11), nrow=2, byrow = TRUE))
par(mar=c(0,2,1,2.5), oma=c(1,1,0,0),xpd=TRUE)
# --------------- Baseline model
dupdat<-data.table::setDT(baseline_points)[,list(Count=.N),names(baseline_points)]
# Calculate the percentage for easier point plotting
dupdat$Percent<-dupdat$Count/sum(dupdat$Count)

# Make the ternary diagram
morphs_ternary_plot(lab.cex=1.5,
                      lab.offset=0.25,
                      atip.pos=3,
                      atip.rotate=0,
                      btip.pos=2, 
                      btip.rotate=0,
                      ctip.pos=2,
                      ctip.rotate=0)

# add the points
AddToTernary(
  points,
  col="black",
  dupdat[,1:3],
  lwd=2,
  cex=10*dupdat$Percent+1
)

#high div point
AddToTernary(
  points,
  col=divcols[1],
  baseline_points[high_which,],
  lwd=2,
  pch=8,
  cex=2
)
#low div point
AddToTernary(
  points,
  col=divcols[2],
  baseline_points[low_which,],
  lwd=2,
  pch=8,
  cex=2
)

# -------------- Single locus sims: low
mar=c(0,2,1,1.5)

lowdat<-as.data.frame(do.call(rbind,lowDiv[low_patts=="monogamy_nm"]))
colnames(lowdat)<-c("CP","NP","NN")
dupdat<-data.table::setDT(lowdat)[,list(Count=.N),
                                  names(lowdat)]
# Calculate the percentage for easier point plotting
dupdat$Percent<-dupdat$Count/sum(dupdat$Count)

morphs_ternary_plot(lab.cex=1.5,
                      lab.offset=0.25,
                      atip.pos=3,
                      atip.rotate=0,
                      btip.pos=2, 
                      btip.rotate=0,
                      ctip.pos=2,
                      ctip.rotate=0)

# add the points
AddToTernary(
  points,
  dupdat[,1:3],
  col=divcols[2],
  pch=1,
  lwd=2,
  cex=10*dupdat$Percent
)


 text(x=0,
       y=1.2,
       "Baseline",
       cex=2,
       xpd=TRUE
       )
mtext("Low diversity",
      2,
      outer=FALSE,
      line=2,
      col=divcols[2],
      cex=1.5)


# -------------- low diversity
  arch_pch<-c(rep(2,3),6)
names(arch_pch)<-names(points_list)

arch_labels<-c(qtls="QTLs",
              prop0.05="Supergene 5%",
              prop0.25="Supergene 25%",
              prop0.5="Supergene 50%")


low_freqs<-final_freqs[which(final_freqs$diversity=="low"),]
points_list_low<-by(low_freqs, low_freqs$arch, function(arch_dat){
  # convert the data into the list format
  points<-apply(arch_dat,
               1,
               function(dat){
                 out<-dat[c("CP","P","NON")]
                 return(as.numeric(out))
                 },
               simplify = FALSE)
  # output the data with the other info
  return(list(points=points,
              chroms=as.factor(arch_dat$nchrom),
              qtls=as.numeric(arch_dat$nqtl),
              arch=unique(arch_dat$arch)
              )
         )
}
)
tmp<-lapply(points_list_low[c("qtls","prop0.05","prop0.25","prop0.5")], function(points_info){
  morphs_ternary_plot(lab.cex=1.5,
                      lab.offset=0.25,
                      atip.pos=3,
                      atip.rotate=0,
                      btip.pos=2, 
                      btip.rotate=0,
                      ctip.pos=2,
                      ctip.rotate=0)
  
  # add the points
  AddToTernary(
    points,
    points_info$points,
    col=alpha(chr_col[points_info$chroms],0.75),
    pch=arch_pch[points_info$arch],
    lwd=2,
    cex=log(points_info$qtls,base = 2)-1
  )
  text(x=0,
       y=1.2,
       arch_labels[points_info$arch],
       cex=2,
       xpd=TRUE
       )
})





# -------------- Single locus sims: high
par(mar=c(2,2,0,1.5))



highdat<-as.data.frame(do.call(rbind,highDiv[high_patts=="monogamy_nm"]))
colnames(highdat)<-c("CP","NP","NN")
dupdat<-data.table::setDT(highdat)[,list(Count=.N),
                                  names(highdat)]
dupdat$Percent<-dupdat$Count/sum(dupdat$Count)

morphs_ternary_plot(lab.cex=1.5,
                      lab.offset=0.25,
                      atip.pos=3,
                      atip.rotate=0,
                      btip.pos=2, 
                      btip.rotate=0,
                      ctip.pos=2,
                      ctip.rotate=0)

AddToTernary(
  points,
  dupdat[,1:3],
  col=divcols[1],
  pch=1,
  lwd=2,
  cex=100*dupdat$Percent # add color based on param values
)

mtext("High diversity",
      2,
      outer=FALSE,
      line=2,
      col=divcols[1],
      cex=1.5)

# --------------- high diversity

tmp<-lapply(points_list[c("qtls","prop0.05","prop0.25","prop0.5")], function(points_info){
  morphs_ternary_plot(lab.cex=1.5,
                      lab.offset=0.25,
                      atip.pos=3,
                      atip.rotate=0,
                      btip.pos=2, 
                      btip.rotate=0,
                      ctip.pos=2,
                      ctip.rotate=0)
  
  # add the points
  AddToTernary(
    points,
    points_info$points,
    col=alpha(chr_col[points_info$chroms],0.75),
    pch=arch_pch[points_info$arch],
    lwd=2,
    cex=log(points_info$qtls,base = 2)-1
  )
})


lgd<-outer_legend(x=-1.15,
                  y=-0.9,
                  c("# Chrom:", names(chr_col)),
                  col=c(NA,chr_col),
                  pch=c(NA,rep(15,3)),
                  cex=2,
                  bty='n',
                  horiz=TRUE,
                  pt.lwd=2,
                 text.width=c(0.2,rep(0.1,3)))
outer_legend(x=0,
             y=-0.9,
             legend=c("# QTLs:", levels(as.factor(unique(high_freqs$nqtl)))),
             pch=c(NA,rep(0,4)),
             col=c(NA,rep("black",4)),
             cex=2,
             pt.cex=c(0,log(as.numeric((levels(as.factor(unique(high_freqs$nqtl))))),base=2)-1),
             horiz=2,
             pt.lwd=2,
             bty='n',
             text.width=c(0.2, rep(0.1,4)),
             inset=0)


```

\elandscape
