---
title: "Ternary plots for ARTs results"
output:
  pdf_document: default
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,out.extra='',fig.pos="H",dpi=200,fig.height = 7,fig.width = 7)
knitr::opts_knit$set(root.dir='../results/') #change
```
```{r opts, echo = FALSE}
library(knitr)
knitr::opts_chunk$set(
  fig.path = "../figs/",
  dev="png"
)
```
```{r librarysetup, echo=FALSE,echo=FALSE,message=FALSE}
library(RColorBrewer)
library(scales)
library(sm)
library(kableExtra)
library(plotly)
library(dplyr)
library(vegan)
library(tidyr)
library(Ternary)
source("../R/002_freq_functions.R")
tmp<-sapply(list.files(path="../../spfTools/R/",full.names = TRUE),source)
source("../morph_predictions/check_freqs.R")
source("../morph_predictions/morph_gens_ns.R")
source("../R/formatting.R")
```


```{r dataConversionFXN}
morphs_to_ternary<-function(pattern, path){
  data<-get.morph.freqs(plot.pc.reps(pattern=pattern,path=path,cols,make.plot=FALSE))
  data_points<-apply(data,
                   1,
                   function(dat){
                     out<-dat[c("FreqCP","FreqNcP","FreqNcNp")]
                     return(as.numeric(out))
                   },
                   simplify = FALSE)
  return(data_points)
}

```
```{r makeTernaryMorphsFXN}
# made with inspiration from this tutorial: https://cran.r-project.org/web/packages/Ternary/vignettes/Ternary.html
morphs_ternary_plot<-function(){
  TernaryPlot(atip="CP",
            btip="NP",
            ctip="NN",
            alab="Courter/Parent",
            blab="Noncourter/Parent",
            clab="Noncourter/Nonparent",
            grid.lines=4,
            grid.minor.lines = 1,
            grid.lty = "dotted",
            grid.minor.lty = "dotted",
            lab.col=cols2[c(1,2,4)],
            tip.col=cols2[c(1,2,4)])
}
```

## Baseline model

The baseline model results are saved in `morph_results_Ns.RDS`, I believe.

```{r baselineData}
morph_results<-readRDS("../morph_predictions/morph_results_Ns.RDS")
morph_results$diversity<-vegan::diversity(round(
  morph_results[,c("CP","CN","NP","NN")],4))

baseline_points<- morph_results[which(morph_results$initial_CP==0.25 &
                                        morph_results$initial_NP==0.25 & 
                                        morph_results$initial_NN==0.25),
                                c("CP","NP","NN")]
baseline_points<-baseline_points[which(rowSums(baseline_points)>0),]

```

```{r baselineTernary}
morphs_ternary_plot()

# add the points
AddToTernary(
  points,
  baseline_points,
  col=alpha("grey",0.5),
  lwd=2,
  cex=2 # add color based on param values
)
```

Note: would be good to scale points based on number of times they land at one outcome...

```{r}
dupdat<-data.table::setDT(baseline_points)[,list(Count=.N),names(baseline_points)]
# Calculate the percentage for easier point plotting
dupdat$Percent<-dupdat$Count/sum(dupdat$Count)

# Make the ternary diagram
morphs_ternary_plot()

# add the points
AddToTernary(
  points,
  col="purple",
  dupdat[,1:3],
  lwd=2,
  cex=5*dupdat$Percent
)
```


## Single locus simulations

These simulations had three diversity settings but also four mating systems, so I'll use a mix of colors and points to show the differences.

```{r singleLocusData}
lowDiv<-morphs_to_ternary(pattern="lowDiversity.*summary.txt",path="single_locus")
highDiv<-morphs_to_ternary(pattern="highDiversity_.*summary.txt", path="single_locus")
highDivStrict<-morphs_to_ternary(pattern="highDiversityStrict.*summary.txt",path="single_locus")

# remove ones with no viability sel or random mating
lowDiv<-lowDiv[grep("v0",grep("RM", names(lowDiv), invert=TRUE, value=TRUE), invert=TRUE, value=TRUE)] 
highDiv<-highDiv[grep("v0",grep("RM", names(highDiv), invert=TRUE, value=TRUE), invert=TRUE, value=TRUE)]
highDivStrict<-highDivStrict[grep("v0",grep("RM", names(highDivStrict), invert=TRUE, value=TRUE), invert=TRUE, value=TRUE)]
```

```{r singleLocusMSinfo}
# extract corresponding mating system info
low_patts<-gsub(".*lowDiversity_(.*)_\\d_summary.txt_\\d","\\1",names(lowDiv))
low_patts<-grep("v0",grep("RM", low_patts, invert=TRUE, value=TRUE), invert=TRUE, value=TRUE)

high_patts<-gsub(".*highDiversity_(.*)_\\d_summary.txt_\\d","\\1",names(highDiv))
high_patts<-grep("v0",grep("RM", high_patts, invert=TRUE, value=TRUE), invert=TRUE, value=TRUE)

strict_patts<-gsub(".*highDiversityStrict_(.*)_\\d_summary.txt_\\d","\\1",names(highDivStrict))
strict_patts<-grep("v0",grep("RM", strict_patts, invert=TRUE, value=TRUE), invert=TRUE, value=TRUE)

```


```{r singleLocusPlot}

ms_cols<-c(monogamy='#e08214',
           monogamy_nm='#b35806',
           polygyny='#998ec3',
           polygyny_nm='#5e3c99')

# create the base
par(mar=c(0,0,0,0))
morphs_ternary_plot()

# add the points
AddToTernary(
  points,
  lowDiv,
  col=ms_cols[low_patts],
  pch=6,
  lwd=2,
  cex=2 # add color based on param values
)

AddToTernary(
  points,
  highDiv,
  col=ms_cols[high_patts],
  pch=0,
  lwd=2,
  cex=2 # add color based on param values
)

AddToTernary(
  points,
  highDivStrict,
  col=ms_cols[strict_patts],
  pch=5,
  lwd=2,
  cex=2 # add color based on param values
)

outer_legend("topleft",
       c(names(ms_cols),"low", "high", "strict high"),
       col=c(ms_cols, rep("dark grey",3)),
       pch=c(rep(1,4),6,0,5),
       pt.cex=2,
       bty='n',
       ncol=2,
       pt.lwd=2)


```


```{r singleLocusSidBySide}


# create the base
par(mar=c(0,0,0,0),
    mfrow=c(1,4))
morphs_ternary_plot()

# add the points
AddToTernary(
  points,
  lowDiv,
  col=ms_cols[low_patts],
  pch=6,
  lwd=2,
  cex=2 # add color based on param values
)

morphs_ternary_plot()
AddToTernary(
  points,
  highDiv,
  col=ms_cols[high_patts],
  pch=0,
  lwd=2,
  cex=2 # add color based on param values
)

morphs_ternary_plot()
AddToTernary(
  points,
  highDivStrict,
  col=ms_cols[strict_patts],
  pch=5,
  lwd=2,
  cex=2 # add color based on param values
)

outer_legend("topleft",
       c(names(ms_cols),"low", "high", "strict high"),
       col=c(ms_cols, rep("dark grey",3)),
       pch=c(rep(1,4),6,0,5),
       pt.cex=2,
       bty='n',
       ncol=2,
       pt.lwd=2)


```

```{r monogamyPlot}

ms_cols<-c(monogamy='#e08214',
           monogamy_nm='#b35806',
           polygyny='#998ec3',
           polygyny_nm='#5e3c99')

# create the base
par(mar=c(0,0,0,0))
morphs_ternary_plot()

# add the points
AddToTernary(
  points,
  lowDiv,
  col=ms_cols[low_patts],
  pch=6,
  lwd=2,
  cex=2 # add color based on param values
)

AddToTernary(
  points,
  highDiv,
  col=ms_cols[high_patts],
  pch=0,
  lwd=2,
  cex=2 # add color based on param values
)

AddToTernary(
  points,
  highDivStrict,
  col=ms_cols[strict_patts],
  pch=5,
  lwd=2,
  cex=2 # add color based on param values
)

outer_legend("topleft",
       c(names(ms_cols),"low", "high", "strict high"),
       col=c(ms_cols, rep("dark grey",3)),
       pch=c(rep(1,4),6,0,5),
       pt.cex=2,
       bty='n',
       ncol=2,
       pt.lwd=2)


```

## Polygenic inheritance

The main variables I want to show with the polygenic inheritance are the genetic architecture (QTLs vs supergenes with diff props) and number of chromosomes. I could also do number of QTLs -- that wasn't shown in the main manuscript, as I showed representative images from the 8 QTLs scenarios. How do I want to show this? 

- In other figures, I've used circles and triangles for QTLs and supergenes, respectively
- I could use colors to show num of chromosomes and do different triangles for each architecture
- scale points based on the number of QTLs?

```{r}
final_freqs<-read.csv("../results/morph_freqs_summary.csv")
final_freqs$rep<-gsub("^.*Diversity_(.*)_pheno.csv","\\1",final_freqs$file)

# extract the relevant info for plotting
final_freqs$nqtl<-as.numeric(gsub("^.*q(\\d+)_.*$","\\1",final_freqs$rep))
final_freqs$nchrom<-as.factor(gsub("^.*c(\\d+)_.*$","\\1",final_freqs$rep))
final_freqs$arch<-gsub("^.*results\\/(\\w+)\\/.*$","\\1",final_freqs$file)
final_freqs$arch[final_freqs$arch=="supergene"]<-gsub("^.*(prop\\d+\\.\\d+).*$","\\1",final_freqs$file[final_freqs$arch=="supergene"])
final_freqs$diversity<-gsub("^.*\\/(\\w+)Diversity.*$","\\1",final_freqs$file)

# only keep the high diversity ones
high_freqs<-final_freqs[which(final_freqs$diversity=="high"),]

```
```{r}
chr_col<-c(
  "#cccccc",
  "#969696",
  "#636363"
)
names(chr_col)<-levels(final_freqs$nchrom)
```


I'll start with just the QTL options.

```{r}

points_list<-by(high_freqs, high_freqs$arch, function(arch_dat){
  # convert the data into the list format
  points<-apply(arch_dat,
               1,
               function(dat){
                 out<-dat[c("CP","P","NON")]
                 return(as.numeric(out))
                 },
               simplify = FALSE)
  # output the data with the other info
  return(list(points=points,
              chroms=as.factor(arch_dat$nchrom),
              qtls=as.numeric(arch_dat$nqtl),
              arch=unique(arch_dat$arch)
              )
         )
}
)




```


```{r architectureTernary}
arch_pch<-c(rep(2,3),6)
names(arch_pch)<-names(points_list)


par(mar=c(0,0,0,0),
    mfrow=c(1,4))

tmp<-lapply(points_list, function(points_info){
  morphs_ternary_plot()
  
  # add the points
  AddToTernary(
    points,
    points_info$points,
    col=alpha(chr_col[points_info$chroms],0.75),
    pch=arch_pch[points_info$arch],
    lwd=2,
    cex=log(points_info$qtls,base = 2)-1
  )
  text(x=0,
       y=1.1,
       points_info$arch,
       cex=2,
       xpd=TRUE
       )
})



lgd<-outer_legend(x=-0.35,
                  y=0.65,
                  names(chr_col),
                  col=chr_col,
                  pch=15,
                  cex=2,
                  bty='n',
                  horiz=TRUE,
                  pt.lwd=2,
                  title="Number of Chromosomes")
outer_legend(x=-0.35,
             y=-0.45,
             legend=levels(as.factor(unique(high_freqs$nqtl))),
             pch=15,
             cex=2,
             pt.cex=log(as.numeric((levels(as.factor(unique(high_freqs$nqtl))))),base=2)-1,
             horiz=2,
             pt.lwd=2,
             bty='n',
             title="Number of QTLs")
```


```{r}
morphs_ternary_plot()

# add the points
AddToTernary(
  points,
  lwd=2,
)


```


















